### 2018-05-17 paper/figpaper.tex
new paper for figures

\documentclass[11pt]{article}
\usepackage{amsmath}

\begin{document}
In order to compare the invariants visually, we use the following
polynomial in two variables:
\begin{equation}
  \begin{split}
    f(x, y) &= 0.5 - 0.2x - 0.3y - 0.05x^2 + 0.03xy + 0.04y^2 \\ 
    &+ 0.03y^3 + 0.001x^3 - 0.0015x^2y + 0.002xy^2 - 0.0005y^3.
  \end{split}
\end{equation}
The function was chosen to have no critical points, and to be predominantly
linear, so that the higher order derivatives were not too large.

\begin{figure}

\end{figure}

\end{document}