### 2019-11-06 paper/tmp.tex
clipboard

Using the same function as our image.

For $A(2)$ the weight of the partial transvectants prevents them from being invariants. They can be normalised by choosing only those where all components have the same weight and projecting them onto a unit hypersphere. Instead, we select ratios of partial transvectants. It is beneficial to choose the numerator and denominator so that they have different critical points, which can be achieved by adding together partial transvectants of the same weight as appropriate.