### 2021-02-16 paper/newpaper.tex
Updates

\documentclass[review,onefignum,onetabnum]{siamonline190516}
\usepackage{amsmath,amsfonts,graphicx}
\def\R{\mathbb{R}}
\usepackage{mathtools}
\usepackage{xcolor}
\usepackage{todonotes}
\usepackage{tikz}
\usepackage{array}
\usepackage{natbib}

\date{\today}
\title{Differential invariant signatures for planar Lie group transformations of images}
\author{Richard Brown, Stephen Marsland, and Robert McLachlan}

\begin{document}
\maketitle
\begin{abstract}
Changes in viewpoint, such as camera motion, impose transformations on the objects in an image, and these transformations will differ for objects at different positions relative to the camera. Typically these transformations will be elements of planar Lie groups. While there have been significant advances in object recognition using machine learning in recent years, the question of how to recognise an object in an image as its appearance varies through camera motion and similar effects remains open. We demonstrate how differential invariant signatures can be derived for each of the transformation groups, and how the underlying invariances reflect the group-subgroup structure of the transformation groups.

There are a variety of methods that can be used to identify differential invariants, and we provide examples of three of them: tensor contraction, transvectants, and the method of moving frames. We use the resulting invariants to construct practical sets to form three-dimensional invariant signatures. These signatures are not necessarily complete: the image cannot always be reconstructed uniquely up to transformation, but they are plottable, and depend at worst on third derivatives, although more channels of information, such as colour images, can reduce the highest order of derivative needed in some cases. We demonstrate the invariant signatures for each transformation group based on a simple smooth image. A full consideration of how these signatures could be used in practice will require effective methods to numerically approximate derivatives for images.
\end{abstract}

% REQUIRED
\begin{keywords}
Object recognition, planar transformation, planar Lie groups, invariants, differential invariant signature
\end{keywords}

% REQUIRED
\begin{AMS}
65D18, 53A55, 68U10
\end{AMS}

\section{Introduction}
The appearance of a fixed object in a set of images can vary markedly depending upon such variables as pose and position relative to the camera, even though the object itself does not change. This is even more marked in scenes that consist of multiple objects at different distances and orientations to the camera against a fixed background. The appearance of each individual object will then vary independently as a constant camera motion or other such transformation is applied. Typically the individual transformations will appear as the action of an element of a planar Lie group, for example the affine, or projective groups, or an angle-preserving transformation from the conformal group, depending upon the camera model.

For image recognition it is then natural to seek representations of objects that do not change as these various transformation groups act. If the space of objects is denoted $M$ and the appearance transformations form a group $\mathcal{G}$, then mathematically one seeks a way to study objects in $M / \mathcal{G}$. One way to achieve this is by transforming all objects to match a chosen target object by registration (the transformation of the appearance of one shape to match another, typically by gradient descent on a matching function, see e.g.,~\citep{Modersitzki03}). For example, for the group of similarity transformations (rotation and reflection, translation, and scale), a Procrustes alignment~\citep{Kendall1989} will place all of the shapes into the reference frame of one template example. However, analysis on these quotient spaces is rarely simple (see e.g.,~\citet{RiemannianStats} for a basic primer), and if one is interested in the object, and not the transformation, then it necessitates a lot of computational work for relatively little benefit. The alternative approach, of showing examples of each object after applying sufficiently many transformations that the classifier can learn to recognise the object after any group transformation, inevitably scales extremely badly.

An alternative option is to seek $\mathcal{G}$-invariant functions on $M$. Mathematical invariants provide a useful way to identify sets of objects that are the same up to some equivalence relation, since the functions are constant for objects in the same equivalence class (orbits of the group). For images, rather than considering the image of the object, one constructs an invariant signature of some kind, which represents the object modulo the action of elements from the relevant group. This has been considered extensively by Olver and co-authors for the case of outlines of objects~\citep{Calabi1998,Hoff2013}. We extend this by considering the case of signatures for images rather than outlines of curves under a variety of transformation groups of relevance to image analysis.

A greyscale image can be viewed as a two-dimensional manifold embedded in a three dimensional ambient space, with coordinates $(x, y, f(x,y))$. A signature $\mathcal{I}(f) \in \mathbb{R}^m$ has coordinates $\mathcal{I}(f) = (I_1, I_2, \ldots, I_m)$, where each $I_i$ is a function of $f(x, y))$ and its derivatives, and will therefore be a two-dimensional manifold embedded in $\mathbb{R}^m$. Under a change of coordinates $(x, y) \mapsto (\bar{x}, \bar{y})$ induced by a group action, the signatures of the original image and transformed version $\bar{f}(\bar{x}, \bar{y}) = f(x, y)$ will be the same. The introduction of an invariant signature reduces the object recognition problem modulo some transformation group to one of comparing the signatures of the objects, rather than the images themselves. We do not consider the question of classification further in this paper; a standard image recognition algorithm can be used (see, e.g.,~\citet{Zhang20} for a survey), or a matching method such as~\citet{UsCurrents} can be used.

Ideally, an invariant signature would be complete, so that the image is exactly determined by the signature up to a transformation, and two objects share a signature if and only if they differ only by a transformation from $\mathcal{G}$. However, there are some pathological scenarios that occur regularly in images, particularly that flat regions of images collapse down to single points in the signature (as all of the derivatives are zero in these regions). Then images that are otherwise similar (apart from the size of one of these regions) cannot be distinguished. However, the number of dimensions $m$ of the ambient manifold can increase markedly for complete signatures, increasing the computational cost and making it impossible to visualise the signature surface. For many applications an incomplete signature is sufficient, and is what we consider here, favouring signatures in three coordinates, since for greyscale images, $\mathbb{R}^3$ is the smallest space in which the signature surfaces can reside.  If the signature is not complete, then there will be a `bad set' of images that cannot be distinguished from each other.  The codimension of the bad set is defined as the difference between the dimensionalities of the whole space and of the bad set. The codimension should ideally be large, meaning that the dimension of the bad set would be small.

In addition to a bad set of high codimension, in order for the invariant signatures to be useful to identify similarity in images, there are two other key considerations: they must vary continuously with respect to the image, and they must be robust with respect to occlusions, where one object obscures some part of another. These constraints are essential if the signature is to be used to detect similarity of images, but the first fails for the rational invariants found in the literature. Such considerations lead one naturally to consider local invariants, i.e., those defined using each point of the object as its own basepoint, together with a neighbourhood in some topology. This is sufficient to compute, e.g., derivatives of an object. For this reason we focus on differential invariants. The disadvantage of such invariants, as we shall discuss, is the need to compute several orders of spatial derivative of the image. See~\citet{UsMobius} and references therein for more discussion about the desirable properties of a set of invariants for image and curve analysis.

In this paper our aim is develop the mathematical underpinnings of constructing differential invariant signatures for images under the action of planar Lie groups, and to summarise the impediments to a practical implementation, including the number of derivatives required. We demonstrate three distinct methods of computing differential invariants: by tensor contraction, using the theory of transvectants, and via the method of moving frames, applying the methods to a variety of the planar Lie groups, providing sample signature sets for the various groups we consider, together with an example of a sample signature surface based on a random transformation of a simple, smooth image.

The set of groups that we consider are shown as a lattice in Fig.~\ref{fig:lattice}, where the size of the groups decrease down the page, and the arrows denote the subgroup relationship.  There is an important relationship between a group and its subgroups for invariants, which is that invariants of a group are automatically invariants of its subgroups. This means that care is needed to distinguish between the two, with further invariants that are known not to be invariant in the supergroup included. The finite-dimensional groups in this lattice were chosen because they are the only planar Lie groups that include translation and have locally primitive algebras (that is, for any open subset of the plane, there is no one-dimensional foliation that is left invariant by the action of the corresponding group~\citep{Gonzalez-Lopez90}). This makes them the groups most directly relevant to image transformations. The infinite-dimensional examples are likewise those most likely to be of interest for image transformation, in particular the full diffeomorphism group, which is commonly used for image registration for medical images, see e.g.,~\citet{YounesBook}.

%We next provide a description of the notation that we use in this paper, followed by a summary of relevant work on invariants. We also introduce the way that we will present examples of the signatures for the various groups we consider. Following this, in the next three sections we describe three distinct methods of computing differential invariants: by tensor contraction, using the theory of transvectants, and via the method of moving frames, applying the methods to a variety of the planar Lie groups. We conclude by considering the practical issues of implementation, including the number of derivatives required and the effect this has on the practicality of the method.

\begin{figure}
\begin{center}
\begin{tikzpicture}[node distance=2cm]
\node(C1) [] {$\mathrm{Diff}$};
\node(C2) [below left of=C1] {$\mathrm{Diff}_{\mathrm{vol}}$};
\node(C3) [below right of=C1] {$\mathrm{Diff}_{\mathrm{con}}$};
\node(B1)    [below of=C1]    {$PSL(3,\mathbb{R})$};
\node(A1) [below of=C3]                          {$PSL(2,\mathbb{C})$};
\node(B2)   [below of=B1]              {$A(2)$};
\node(A2)       [below right of=B2]          {Sim$(2)$};
%\node(A3)      [right of=A2]              {$PSL(2,\mathbb{R})$};
\node(B3)   [below left of=B2]            {$SA(2)$};
\node(A4)      [below of=A2]            {$E(2)$};
\node(A6)      [below left of=A4]            {$SE(2)$};
\node(A5) [below of=A6]{$T(2)$};

\draw[thick,->] (C1) to  (C2);
\draw[thick,->] (C1) to  (C3);
\draw[thick,->] (C1) to  (B1);
\draw[thick,->] (C2) to  (B3);
%\draw[thick,->] (C1) to  (A2);
\draw[thick,->] (C3) to  (A1);
\draw[thick,->] (A1)   to (A2);
\draw[thick,->] (B1) to  (B2);
\draw[thick,->] (B2) to  (B3);
\draw[thick,->] (B2) to  (A2);
\draw[thick,->] (A2) to  (A4);
\draw[thick,->] (B3) to  (A6);
\draw[thick,->] (A4) to  (A6);
\draw[thick,->] (A6) to  (A5);
\end{tikzpicture}
\end{center}
\caption{The lattice of groups considered in this paper. The edges identify subgroups.}
\label{fig:lattice}
\end{figure}

\section{Preliminaries}

\subsection{Concepts and Notation}

We define a {\em $k$-colour image} as a triple $(f,\Omega,k)$ where $\Omega\subset\R^n$ (typically, images are planar, so $n=2$), $1\le k\in\mathbb{Z}$ is the number of colour channels, and $f\colon\Omega\to\R^k$. The function $f$ will be taken to be as smooth as necessary. Consider a (finite or infinite-dimensional) local transformation group $\mathcal{G}$.  A transformation $\varphi\in G$ acts on images by $\varphi\cdot (f,\Omega,k) = (f\circ\varphi^{-1},\varphi(\Omega),k)$. We mostly consider greyscale images, so $k=1$. Also, since we work locally, we will usually omit the domain $\Omega$.

In order to define differential invariants, it is necessary to identify how the group transformation $\varphi$ acts on derivatives of the function $f$ via the jet space~\citep{OlverEIS}. The $d$-th order jet space $J^d$ consists of the function itself and the set of all partial derivatives of the function up to order $d$. (Note that there are $\left( \begin{array}{c} d+k-1 \\ k \end{array} \right)$ partial derivatives of order $d$.) As the group $\mathcal{G}$ is local, it acts naturally on derivatives of the transformation, and so each individual transformation $\varphi \in \mathcal{G}$ can be prolonged to act on the jet space $J^d$. We can thus prolong the action of the group and consider the action of $\mathcal{G}$ on not just $f$, but all derivatives of $f$ up to order $d$. A $d$-th order differential invariant is then a local scalar function $I: J^{d} (f,\Omega, k) \to \mathbb{R}$ that is invariant under the action of $\mathcal{G}$, i.e., $I(\varphi(f,\Omega, k)) = I(f, \Omega, k) \forall \varphi \in \mathcal{G}$. It can also be useful to consider the maximum order of derivative in the invariant, which we term degree, e.g., $f_{xy}^2 f_x f_y$ has degree 2.

An signature set is the image of $m$ invariants
$\mathcal{I}=(I_1,I_2,\dots,I_m)$. For simplicity of notation we will write
$\mathcal{I}(f) \subset \mathbb{R}^m$ with the understanding that each signature component is
actually acting on the jet space $J^d(f, \Omega, k)$. See Section~\ref{sec:T2} for a basic
example. A signature set is complete for $f$ if it determines $f$ up to
transformations from $\mathcal{G}$, i.e., if $\mathcal{I}(\tilde
f)(\tilde\Omega) = \mathcal{I}(f)(\Omega)$ implies that there exists a
transformation $\varphi\in \mathcal{G}$ such that $\tilde f = \varphi\cdot f$
and $\tilde \Omega = \varphi(\Omega)$.

The methods we use to construct the invariants, particularly the moving frames, produces invariants that, when considered alone, are singular at points (such as critical points). However, in all groups considered the invariants are rational functions of the derivatives of the function. Moreover, the denominators of these functions are powers of each other, and so it is possible to clear them by a projection technique that will be introduced in Section~\ref{sec:Sim2}.

For the case of two-dimensional greyscale images it is often convenient to work in coordinates. In this case the transformation group action by a coordinate transformation $\varphi \colon \mathbb{R}^2 \to \mathbb{R}^2$ is given by (where $f$ and $\mathbf{x}$ denote the original image and coordinates, and $\bar{f}$ and $\bar{\mathbf{x}}$ the transformed versions):
\begin{equation}\label{eq:transformation}
  \mathbf{x} \mapsto \bar{\mathbf{x}} = \varphi(\mathbf{x}).
\end{equation}
The transformed image is then defined to be:
\begin{equation}\label{eq:fbarequalsf}
  \bar{f}(\bar{\mathbf{x}}) = (f\circ\varphi^{-1})(\bar{\mathbf{x}}) = f(\mathbf{x}).
\end{equation}

Derivatives of $\bar{f}$ and $f$ can be related by the chain rule, for example the first order derivatives are given by:
\begin{equation}\label{eq:prolongation}
  \begin{aligned}
    f_x &= \bar{f}_{\bar{x}} \bar{x}_x + \bar{f}_{\bar{y}}\bar{y}_x \\
    f_y &= \bar{f}_{\bar{x}} \bar{x}_y + \bar{f}_{\bar{y}}\bar{y}_y \\
  \end{aligned}
\end{equation}
where $\bar{x}_x = \frac{\partial \varphi_1}{\partial x}(x, y)$, et cetera. %It is often convenient to work with $\varphi^{-1}$ instead, as in \S\ref{sec:movingframes}.

%Groups we've done:
%translations, E(2), Sim(2), SA(2), A(2), PSL(2,C), PSL(3,R), Diff\_con, Diff\_vol, Diff
\subsection{Test image}

For each signature set that we derive, it is informative to see an example signature of the kind of surface that is defined, 
normally a two-dimensional surface in three-dimensions.  We will apply a small random transformation from 
the relevant group to a single continuous image and plot the signature before and after this transformation has been applied.
This is an aid to understanding rather than any form of demonstration of the robustness of our signature sets.

The image we use is a continuous function (thus avoiding numerical issues in the approximation 
of derivatives) defined on $[-1,1] \times [-1,1]$ as:

\begin{equation}\label{eq:function}
f(x, y) = \exp(-2x^2 - 4\sin^2(y + 0.5x^2))
\end{equation}

The function is depicted as a greyscale image in Fig.~\ref{fig:function}.
We will use this function throughout, using the grid on the right of the figure to help visualise both the transformation and how the image
is mapped into the three-dimensional signature surface.

\begin{figure}
  \centering
  \includegraphics[width=.45\textwidth]{Figs/function}
  \includegraphics[width=.45\textwidth]{Figs/function_scanlines}
  \caption{The test function used as an image, defined in Eq.~\eqref{eq:function}.
On the left the image itself is shown, while on the right a version with a
regular grid superimposed is given. 
}\label{fig:function}
\end{figure}

\subsection{Getting Started: Translations}
\label{sec:T2}

The most basic transformation group of a image is the translation group
$T(2) \colon \mathbb{R}^2 \to \mathbb{R}^2$ with elements of the form
$\varphi(\mathbf{x}) = \mathbf{x} + \mathbf{t}$ for fixed $\mathbf{t}$.
These could arise from actions such as cropping a photo.

We could explicitly write the calculation in Eq.~\ref{eq:prolongation} for
this case, but instead provide an intuitive explanation. Consider a point
in the original image, and the corresponding point in the transformed one.
Clearly, the intensity value of the points match, and so do their first
spatial derivatives. These three quantities can be used as coordinates in a
new space, where the two image points will correspond. Computing these
coordinates for all image points will define a surface in the new
three-dimensional space, and the two images will produce the same surface,
the signature, $\mathcal{I}(f)(x, y) = \left(I_1, I_2,
I_3\right)$ where $I_1 = f, I_2 = f_x, I_3 = f_y$.

Fig.~\ref{fig:T2} shows (a) the original image, (b) the translated image
under the transformation $(x,y) \mapsto (x + 0.1, y - 0.2)$, (c) the
signature of one of them, (d) the superimposed pair of signatures. The
yellow lines in (c) correspond to places where the yellow grid from (a) has been
mapped to on the signature surface. The parts of the surface where the mesh
gets close together are where the function and its derivatives are tending
towards zero. This occurs around the boundary of our test image and so all
of these points are close together in the signature space. In (d), notice
that as some of the image has moved out of frame, there are parts of the
signature surfaces of the two images that do not overlap.

\begin{figure}
  \centering
  \begin{tabular}{cc}
  \includegraphics[width=.45\textwidth]{Figs/function_scanlines} &
  \includegraphics[width=.45\textwidth]{Figs/f_transformed_T2} \\
  (a) & (b) \\
  \includegraphics[width=.45\textwidth]{Figs/T2_signature} &
  \includegraphics[width=.45\textwidth]{Figs/T2_match} \\
  (c) & (d)
  \end{tabular}
  \caption{(a), (b) The image before and after the $T(2)$ transformation $(x,
  y) \mapsto (x + 0.1, y - 0.2)$. (c) the signature of the
  image (before translation). (d) the signatures before and after the
  translation superimposed.}\label{fig:T2}
\end{figure}

\subsection{Relevant Literature}

The theory of invariants has a rich mathematical history, starting from invariants of forms~\citep{Cayley}, where the invariants are polynomials in the coefficients of the forms, see e.g.,~\citet{OlverCIT,Kraft2002} for an overview. Geometric considerations and differential invariants followed fairly quickly~\citep{Veblen1872,Tresse1894,Bouton1898,Patterson1928}, particularly because they enabled the classification of differential equations into equivalence classes~\citep{Lie1884,Littlewood44, Cartan52}. More recently, this has been extended, primarily by Olver and collaborators, see e.g.,~\citet{OlverCIT} and~\citet{OlverEIS} for summaries.

There has been a long-standing interest in constructing invariants for image recognition and pattern analysis, see e.g.,~\citet{Mundy1992} and~\citet{Wood1996}. This was set in the context of Lie's invariants by~\citet{VanGool1995}, where the concept of a joint (or semi-) differential invariant (an invariant that acts on the joint action of the transformation group, see~\citet{Fels1997,Olver2013}) was introduced~\citep{VanGool1992}; see also~\citet{Hubert2009}.

A variety of function spaces have allowed for the creation of different sets of invariants for particular groups. For the Euclidean group, the Fourier transform provides a useful tool~\citep{Ghorbel1994,Turski2006,Smach2007,Gauthier2008}, and this can be extended to the affine group using the Fourier-Mellin transform~\citep{Reddy1996,Zhang2019} and the bispectrum~\citep{Kakarala2009,Negrinho2013}. Alternatively, integral invariants~\citep{Manay2006,Feng2010} and geometric moments can be computed. For the similarity group applied to images the original reference is \citet{Hu1962}, but see also~\citet{Hickman2011} for affine transformations, \citet{Flusser2009} for an implicit representation, and \citet{Papakostas2013} for their use in a binary pattern for image recognition, which is somewhat related to the invariant features such as SURF~\citep{SURF}.

The group that has been of particular interest for image analysis is the projective group, because of its ubiquity in camera models. Invariants for this group have been computing using points~\citep{Gros1992,Suk2000}, geometric moments~\citep{Suk2004}, curvature~\citep{Hann2002a}, and differential and joint invariants~\citep{Hubert2007,Arora2009,Kogan2014,Li19}.

Although there are computational challenges regarding the numerical approximation of derivatives on images, the fact that differential invariants are robust to occlusion, cheap to compute, and well-behaved with respect to small changes in the image, has made them popular for researchers in image analysis~\citep{Florack1993}. Two particular tools have been key to this: Cartan's moving frame~\citep{Cartan35} (see Section~\ref{sec:movingframes}), and which was used to construct image invariants as early as 1994 by \citet{Faugeras1994}, and the concept of an image signature built from a set of invariants~\citep{Calabi1998}. An algorithm to compute differential and joint invariants and determine complete signature sets is described by \citet{Olver2005,Olver2007}. The entire set of invariants is computed using the method of moving frames and then the signature set of a maximal set of functionally independent invariants is identified. A third step is possible, in which the dimension of the signature set is reduced by eliminating those invariants whose values can be computed from the values of derivatives of some subset of the invariants.

%Groups we've done:
%translations, E(2), Sim(2), SA(2), A(2), PSL(2,C), PSL(3,R), Diff\_con, Diff\_vol, Diff
\section{Differential invariants through tensor contraction}
\label{sec:contraction}

While we avoid the use of tensors in most of this study, there is a notable
exception to be made for the Euclidean transformation group $E(2)$. In this
setting we can restrict ourselves to \emph{Cartesian tensors}, where we
need make no distinction between covariance and contravariance as they are
equivalent. We therefore use subscript indices exclusively.

In this setting, a Euclidean transformation acting on $\mathbb{R}^2$ is given
by an expression of the form:
\begin{equation*}
  \bar{x}_i = a_{ij}x_j + b_j,
\end{equation*}
using the Einstein summation convention. The orthogonality condition is
that $a_{ik}a_{jk} = a_{ki}a_{kj} = \delta_{ij}$. A Euclidean
transformation $\varphi \colon \mathbb{R}^2 \to \mathbb{R}^2$ can also be
written in matrix form as $\bar{\mathbf{x}} = \varphi(\mathbf{x}) =
U\mathbf{x} + \mathbf{t}$ where the matrix $U$ satisfies $U^TU = I$ and
$\mathbf{t} \in \mathbb{R}^2$.

With respect to $\mathbb{R}^2$, a Cartesian tensor of rank $p$, or $p$-tensor is a $2^p$-tuple of real numbers $C_{i_1\ldots i_p}$ whose components transform under a Euclidean transformation as:
\begin{equation*}
  \bar{C}_{i_1\ldots i_p} = a_{i_1j_1}\cdots a_{i_pj_p} C_{j_1\ldots j_p}
\end{equation*}
and a $0$-tensor is a scalar, or invariant.

The main observation is that in this setting, the $n^\text{th}$ order
partial derivative operator $\partial^n / \partial_{x_{i_1}}\cdots
\partial_{x_{i_n}}$ formally transforms as an $n$-tensor, \citep{Florack1993}. Because the product of a $p$-tensor and a $q$-tensor gives a $(p+q)$-tensor and contraction of a $p$-tensor yields a $(p-2)$-tensor, a complete contraction of a tensor of even rank formed as a product of terms of the form $\partial^p f / \partial_{x_{i_1}} \cdots \partial_{x_{i_p}}$ will give a Euclidean invariant.

For example, up to second derivatives, we have the following Euclidean invariants, expressed as tensor contractions:
\begin{equation}
\label{eq:e2tensor}
\begin{aligned}
  I_0 &= f = f \\
  I_1 &= f_if_i = f_x^2 + f_y^2 \\
  I_2 &= f_{ii} = f_{xx} + f_{yy} \\
  I_3 &= f_{ij} f_{ij} = f_{xx}^2 + 2f_{xy}^2 + f_{yy}^2 \\
  I_4 &= f_i f_j f_{ij} = f_x^2 f_{xx} + 2f_x f_y f_{xy} + f_y^2 f_{yy}.
\end{aligned}
\end{equation}

While we can construct any number of invariants this way, there is no
guarantee of functional independence between them. We will address this
issue in Section~\ref{sec:movingframes} when we construct invariants of
$E(2)$ and other groups using the method of moving frames. We can form a
signature from any three of the invariants in Eq.~\eqref{eq:e2tensor}, for simplicity we choose 
our signature to be:
\begin{equation}
    \label{eq:e2signature}
    \mathcal{I}_{E(2)} = (I_0, I_1, I_2) .
\end{equation}  
A sample $E(2)$ transformation is depicted in Fig.~\ref{fig:E2} with its corresponding
signature.
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_E2.png}
\includegraphics[width=.45\textwidth]{Figs/E2_signature.png}
\caption{A sample $E(2)$ Transform and 3D signature.}
\label{fig:E2}
\end{figure}
We will defer discussion of the Special Euclidean group (which excludes
reflections) until Section~\ref{sec:SE2}, where suitable invariants naturally arise
from the method of moving frames.

\subsection{The similarity group}
\label{sec:Sim2}
The similarity group $\text{Sim}(2)$ consists of transformations of the form
$\bar{\mathbf{x}} = \varphi(\mathbf{x}) = sU\mathbf{x} + \mathbf{t}$, where $U^TU = I$ and $s
\in \mathbb{R}\setminus\{0\}$. It is like the Euclidean group except it
also allows for isotropic scaling. In image applications, similarity
transformations arise from operations like zooming.

Under similarity transformations, the Euclidean invariants become
\emph{relative invariants}, i.e., each derivative introduces a multiple of $s$ to
each invariant. For example, $\bar{f}_{\bar x}^2 + \bar{f}_{\bar y}^2 =
s^2(f_x^2 + f_y^2)$ and $\bar{f}_{\bar x \bar x}^2 + 2\bar{f}_{\bar x \bar
y}^2 + \bar{f}_{\bar y \bar y} = s^4 (f_{xx}^2 + 2f_{xy}^2 + f_{yy}^2$). We
refer to the power of $s$ as the weight of the transformation. The concept of a 
relative invariant will be important in the rest of this paper. It follows the 
definition given here, except that the function weighting the invariant is more 
general, see~\citet{Fels1997} for more details. In contrast to that paper, 
we will continue to use the term weight to represent the power to which the function is raised.

One way to produce similarity invariants would be to take ratios of powers
of the Euclidean invariants \eqref{eq:e2tensor} so that the multiples of
$s$ cancel. For example $\frac{f_x^2 + f_y^2}{f_{xx} + f_{yy}}$ is the
ratio of two weight-2 relative invariants and is hence a similarity
invariant. However, this invariant is singular whenever $f_{xx} + f_{yy} =
0$, which is a codimension 0 phenomenon for images.

A better approach is to take three invariants of the same weight.
Taking $I_1^2$, $I_2^2$, and $I_3$ from \eqref{eq:e2tensor}, all of which are
weight $4$, they can be projected on to the unit sphere, which can be
multiplied by $I_0 = f$ to form a signature surface. This signature is
defined by:
\begin{equation}
    \label{eq:sim2signature}
\mathcal{I}_{\text{Sim(2)}}(f) = \frac{f}{\sqrt{I_1^4 + I_2^4 + I_3^2}}(I_1^2, I_2^2,
        I_3) .
\end{equation}
This signature is only singular when $f_x = f_y = f_{xx} = f_{yy} = f_{xy}
= 0$. A sample similarity transformation and computed signature is shown in
Fig.~\ref{fig:Sim2}.

\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_Sim2.png}
\includegraphics[width=.45\textwidth]{Figs/Sim2_signature.png}
\caption{A sample $\text{Sim}(2)$ Transform and 3D signature.}
\label{fig:Sim2}
\end{figure}

\section{Differential invariants through transvectants}

Another way to compute invariants is to compare the simultaneous action of a group element on separate copies of the underlying space (for planar images, $\mathbb{R}^2$). By identifying these spaces with each other it is possible to construct a function known as a transvectant that is a relative invariant to that group action.

We will demonstrate this for the affine group $A(2)$ and its subgroup
$SA(2)$ where, following the treatment by~\citet{OlverCIT}, this leads to
Cayley's Omega Process~\citep{Cayley1846}.

Consider $\mathbf{x}_i \in \mathbb{R}^2, i=1,2$, so that $\mathbf{x}_i$ can
be written in coordinates as $(x_i, y_i)$; a pair of such $\mathbf{x}$ lie in $\mathbb{R}^2 \times \mathbb{R}^2$. Now consider applying the same affine transformation to each $\mathbf{x}$ independently: $\mathbf{x}_i \mapsto A \mathbf{x}_i + \mathbf{b}$ where $A \in GL(2,\mathbb{R}$) and $\mathbf{b} \in \mathbb{R}^2$. Then the Omega Process is a second-order differential operator defined by: 
\begin{equation}
\Omega_{ij} = \begin{vmatrix} \frac{\partial}{\partial x_i} &
\frac{\partial}{\partial y_i} \\ \frac{\partial}{\partial x_j} &
\frac{\partial}{\partial y_j} \end{vmatrix} = 
\frac{\partial^2}{\partial x_i
\partial y_j} - \frac{\partial^2}{\partial x_j \partial y_i}.
\end{equation}

Under the above affine transformation, $\Omega_{ij} \mapsto (\det A)^{-1}
\Omega_{ij}$, and is therefore a relative invariant. Note that $(\det A)^{-1}$ has no dependence on the coordinates.
The Omega Process can be applied to products of pairs of smooth functions as follows,
writing $\frac{\partial f}{\partial x}$ as $f_x$:

\begin{equation}
\Omega_{ij} \left(f(x_i, y_i) g(x_j, y_j)\right) = f_{x_i}g_{y_j} - f_{y_i}g_{x_j}.
\end{equation}

Identifying $x_i = x_j = x, y_i = y_j = y$ gives functions of $x$ and $y$
that only vary by a scaling factor of $(\det A)^{-1}$ when $\mathbf{x}$ is
mapped by an affine transformation. This matches the definition of a
first-order partial transvectant of two functions:

\begin{equation}
\mbox{tr} \Omega_{ij} f(x_i, y_i) g(x_j, y_j),
\end{equation}

\noindent where $\mbox{tr}$  is the operator that identifies the coordinates of each function $x_i = x_j = x, y_i = y_j = y$.

This can be generalised to $n$ functions $f^{(1)}(x_1, y_1), \ldots, f^{(n)}(x_n, y_n)$ and $r$-th order~\citep{OlverEIS} to define a partial transvectant as:

\begin{equation}
\mbox{tr} \left[\left( \Pi_{k=1}^r \Omega_{i_k j_k} \right) f^{(1)}(x_1, y_1)
f^{(2)}(x_2, y_2), \ldots f^{(n)} (x_n, y_n)\right],
\end{equation}
where $i_k \neq j_k \in \{1, \ldots, n\}$.

Under this formula there is a factor $(\det A)^{-r}$ in place of $(\det
A)^{-1}$ that we saw previously, when all $n$ copies of $\mathbb{R}^2$ are
transformed by the same affine function $\mathbf{x} \mapsto A\mathbf{x} +
\mathbf{b}$. Note that the individual pairwise Omega processes commute.

For the current case it is sufficient to consider $n$ copies of the same
function $f$. If we represent each copy of $\mathbb{R}^2$ as a node in an
undirected graph, and each Omega process $\Omega_{ij}$ as an edge joining
nodes $i$ and $j$, a partial transvectant can be compactly represented in a
graphical form. For example, consider a weight $2$ transvectant defined as:
\begin{align*}
    \mbox{tr} &\left[(\Omega_{12}\Omega_{12}) f(x_1, y_1)f(x_2, y_2)\right] \\
    &= \mbox{tr}\left[(\partial_{x_1 y_2} - \partial_{x_2y_1})(\partial_{x_1 y_2} -
    \partial_{x_2y_1})f(x_1, y_1)f(x_2, y_2)\right]\\
    &= \mbox{tr}\left[\left(\partial_{x_1 y_2} - \partial_{x_2 y_1}\right)\left(f_x(x_1, y_1)f_y(x_2, y_2) - f_y(x_1, y_1)f_x(x_2, y_2)\right)\right] \\
    &= \mbox{tr} 2 \left[f_{xx}(x_1, y_1)f_{yy}(x_2, y_2) - f_{xy}(x_1, y_1)f_{xy}(x_2, y_2)\right] \\
    &= 2 f_{xx}(x, y) f_{yy}(x, y) - 2 f_{xy}(x, y)^2.
\end{align*}
This is a degree $2$ partial transvectant as it involves derivatives of $f$
up to second-order, and can be represented graphically by the following
diagram:
\begin{center}
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.15cm] (1) at (-0.5,0) {1};
    \node[draw,circle,minimum size=0.15cm] (2) at (0.5,0) {2};
    \draw[-] (1) to [out=30,in=150] (2);
    \draw[-] (1) to [out=-30,in=-150] (2);
\end{tikzpicture}
\end{center}
where the double edge corresponds to the two copies of $\Omega_{12}$ (the names are purely for convenience), i.e.,
the weight is the number of edges in the graph.

Tables~\ref{tab:transvectants1} and~\ref{tab:transvectants2} shows all possible non-zero partial
transvectants of $A_2$ up to weight $4$ that can be generated in this way.

\begin{table}
\centering
\begin{tabular}{cccp{9cm}}
Degree & Weight & Diagram & Partial transvectant\\
\hline
2 & 2 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.15cm] (1) at (-0.5,0) {1};
    \node[draw,circle,minimum size=0.15cm] (2) at (0.5,0) {2};
    \draw[-] (1) to [out=30,in=150] (2);
    \draw[-] (1) to [out=-30,in=-150] (2);
\end{tikzpicture}
& $2 f_{xx} f_{yy} - 2 \left(f_{xy}\right)^{2}$ \\

2 & 2 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.15cm] (1) at (0,0.866) {1};
    \node[draw,circle,minimum size=0.15cm] (2) at (-0.5,0) {2};
    \node[draw,circle,minimum size=0.15cm] (3) at (0.5,0) {3};
    \draw[-] (1) to (2);
    \draw[-] (1) to (3);
\end{tikzpicture}
& $\left(f_{x}\right)^{2} f_{yy} - 2 f_{x} f_{y} f_{xy} + f_{xx}
\left(f_{y}\right)^{2}$ \\

2 & 4 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.15cm] (1) at (0,1) {1};
    \node[draw,circle,minimum size=0.15cm] (2) at (0,0) {2};
    \node[draw,circle,minimum size=0.15cm] (3) at (1,0) {3};
    \node[draw,circle,minimum size=0.15cm] (4) at (1,1) {4};
    \draw[-] (1) to (2);
    \draw[-] (2) to (3);
    \draw[-] (3) to (4);
    \draw[-] (4) to (1);
\end{tikzpicture}
& $2 \left(f_{xx}\right)^{2} \left(f_{yy}\right)^{2} - 4 f_{xx} f_{yy} \left(f_{xy}\right)^{2} + 2 \left(f_{xy}\right)^{4}$ \\

2 & 4 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.15cm] (1) at (0,1) {1};
    \node[draw,circle,minimum size=0.15cm] (2) at (-0.9511,0.3090) {2};
    \node[draw,circle,minimum size=0.15cm] (3) at (-0.5878,-.8090) {3};
    \node[draw,circle,minimum size=0.15cm] (4) at (0.5878,-.8090) {4};
    \node[draw,circle,minimum size=0.15cm] (5) at (0.9511,0.3090) {5};
    \draw[-] (1) to (2);
    \draw[-] (2) to (3);
    \draw[-] (3) to (4);
    \draw[-] (4) to (5);
\end{tikzpicture}
& $\left(f_{x}\right)^{2} f_{xx} \left(f_{yy}\right)^{2} - \left(f_{x}\right)^{2} f_{yy} \left(f_{xy}\right)^{2} - 2 f_{x} f_{xx} f_{y} f_{yy} f_{xy} 
 + 2 f_{x} f_{y} \left(f_{xy}\right)^{3} + \left(f_{xx}\right)^{2} \left(f_{y}\right)^{2} f_{yy} - f_{xx} \left(f_{y}\right)^{2} \left(f_{xy}\right)^{2}$ \\

3 & 3 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.15cm] (1) at (0,0.866) {1};
    \node[draw,circle,minimum size=0.15cm] (2) at (-0.5,0) {2};
    \node[draw,circle,minimum size=0.15cm] (3) at (0.5,0) {3};
    \draw[-] (1) to [out=-150,in=90] (2);
    \draw[-] (1) to [out=-90,in=30] (2);
    \draw[-] (1) to (3);
\end{tikzpicture}
& $- f_{x} f_{xx} f_{yyy} - f_{x} f_{yy} f_{xxy} + 2 f_{x} f_{xy} f_{xyy} +
f_{xx} f_{y} f_{xyy} + f_{xxx} f_{y} f_{yy} - 2 f_{y} f_{xy} f_{xxy}$ \\

3 & 3 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.15cm] (1) at (0,1) {1};
    \node[draw,circle,minimum size=0.15cm] (2) at (0,0) {2};
    \node[draw,circle,minimum size=0.15cm] (3) at (1,0) {3};
    \node[draw,circle,minimum size=0.15cm] (4) at (1,1) {4};
    \draw[-] (1) to (2);
    \draw[-] (1) to (3);
    \draw[-] (1) to (4);
\end{tikzpicture}
& $- \left(f_{x}\right)^{3} f_{yyy} + 3 \left(f_{x}\right)^{2} f_{y} f_{xyy} - 3 f_{x} \left(f_{y}\right)^{2} f_{xxy} + f_{xxx} \left(f_{y}\right)^{3}$  \\

3 & 4 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.15cm] (1) at (0,0.866) {1};
    \node[draw,circle,minimum size=0.15cm] (2) at (-0.5,0) {2};
    \node[draw,circle,minimum size=0.15cm] (3) at (0.5,0) {3};
    \draw[-] (1) to [out=-150,in=90] (2);
    \draw[-] (1) to [out=-90,in=30] (2);
    \draw[-] (1) to (3);
    \draw[-] (2) to (3);
\end{tikzpicture}
& $2 f_{xx} f_{yyy} f_{xxy} - 2 f_{xx} \left(f_{xyy}\right)^{2} + 2 f_{xxx}
f_{yy} f_{xyy} - 2 f_{xxx} f_{yyy} f_{xy} - 2 f_{yy}
\left(f_{xxy}\right)^{2} + 2 f_{xy} f_{xyy} f_{xxy}$ \\

%=====

3 & 4 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.15cm] (1) at (0,1) {1};
    \node[draw,circle,minimum size=0.15cm] (2) at (0,0) {2};
    \node[draw,circle,minimum size=0.15cm] (3) at (1,0) {3};
    \node[draw,circle,minimum size=0.15cm] (4) at (1,1) {4};
    \draw[-] (1) to [out=-120,in=120] (2);
    \draw[-] (1) to [out=-60,in=60] (2);
    \draw[-] (2) to (3);
    \draw[-] (3) to (4);
\end{tikzpicture}
& $- f_{x} f_{xx} f_{yy} f_{xyy} + f_{x} f_{xx} f_{yyy} f_{xy} - f_{x} f_{xxx} \left(f_{yy}\right)^{2} + 3 f_{x} f_{yy} f_{xy} f_{xxy} - 2 f_{x} \left(f_{xy}\right)^{2} f_{xyy} - \left(f_{xx}\right)^{2} f_{y} f_{yyy} - f_{xx} f_{y} f_{yy} f_{xxy} + 3 f_{xx} f_{y} f_{xy} f_{xyy} + f_{xxx} f_{y} f_{yy} f_{xy} - 2 f_{y} \left(f_{xy}\right)^{2} f_{xxy}$ \\

3 & 4 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.15cm] (1) at (0,1) {1};
    \node[draw,circle,minimum size=0.15cm] (2) at (0,0) {2};
    \node[draw,circle,minimum size=0.15cm] (3) at (1,0) {3};
    \node[draw,circle,minimum size=0.15cm] (4) at (1,1) {4};
    \draw[-] (1) to [out=-120,in=120] (2);
    \draw[-] (1) to [out=-60,in=60] (2);
    \draw[-] (2) to (3);
    \draw[-] (1) to (4);
\end{tikzpicture}
& $2 \left(f_{x}\right)^{2} f_{yyy} f_{xxy} - 2 \left(f_{x}\right)^{2} \left(f_{xyy}\right)^{2} - 2 f_{x} f_{xxx} f_{y} f_{yyy} + 2 f_{x} f_{y} f_{xyy} f_{xxy} + 2 f_{xxx} \left(f_{y}\right)^{2} f_{xyy} - 2 \left(f_{y}\right)^{2} \left(f_{xxy}\right)^{2}$ \\

% =========
3 & 4 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.15cm] (1) at (0,1) {1};
    \node[draw,circle,minimum size=0.15cm] (2) at (-0.9511,0.3090) {2};
    \node[draw,circle,minimum size=0.15cm] (3) at (-0.5878,-.8090) {3};
    \node[draw,circle,minimum size=0.15cm] (4) at (0.5878,-.8090) {4};
    \node[draw,circle,minimum size=0.15cm] (5) at (0.9511,0.3090) {5};
    \draw[-] (1) to (2);
    \draw[-] (1) to (3);
    \draw[-] (1) to (4);
    \draw[-] (4) to (5);
\end{tikzpicture}
& $- \left(f_{x}\right)^{3} f_{yy} f_{xyy} + \left(f_{x}\right)^{3} f_{yyy} f_{xy} - \left(f_{x}\right)^{2} f_{xx} f_{y} f_{yyy} + 2 \left(f_{x}\right)^{2} f_{y} f_{yy} f_{xxy} - \left(f_{x}\right)^{2} f_{y} f_{xy} f_{xyy} + 2 f_{x} f_{xx} \left(f_{y}\right)^{2} f_{xyy} - f_{x} f_{xxx} \left(f_{y}\right)^{2} f_{yy} - f_{x} \left(f_{y}\right)^{2} f_{xy} f_{xxy} - f_{xx} \left(f_{y}\right)^{3} f_{xxy} + f_{xxx} \left(f_{y}\right)^{3} f_{xy}$ \\
\end{tabular}
\caption{Details and graphical representations of all partial transvectants up to degree $3$ of copies of a function $f$ under a transformation in $A(2)$.}
\label{tab:transvectants1}
\end{table}

\begin{table}
\centering
\begin{tabular}{cccp{9cm}}
Degree & Weight & Diagram & Partial transvectant\\
\hline
4 & 4 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.15cm] (1) at (-0.5,0) {1};
    \node[draw,circle,minimum size=0.15cm] (2) at (0.5,0) {2};
    \draw[-] (1) to [out=30,in=150] (2);
    \draw[-] (1) to [out=-30,in=-150] (2);
    \draw[-] (1) to [out=10,in=170] (2);
    \draw[-] (1) to [out=-10,in=-170] (2);
\end{tikzpicture}
& $f_{xxxx} f_{yyyy} - 4 f_{xyyy} f_{xxxy} + 3 f_{xxyy}^2$ \\

4 & 4 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.15cm] (1) at (0,0.866) {1};
    \node[draw,circle,minimum size=0.15cm] (2) at (-0.5,0) {2};
    \node[draw,circle,minimum size=0.15cm] (3) at (0.5,0) {3};
    \draw[-] (1) to [out=-150,in=90] (2);
    \draw[-] (1) to [out=-90,in=30] (2);
    \draw[-] (1) to [out=-90,in=150] (3);
    \draw[-] (1) to [out=-30,in=90] (3);
\end{tikzpicture}
& $ f_{xx}^2 f_{yyyy} + 2 f_{xx} f_{yy} f_{xxyy} - 4 f_{xx} f_{xy} f_{xyyy} - 4 f_{yy} f_{xy} f_{xxxy} + 4 f_{xy}^2 f_{xxyy} + f_{xxxx} f_{yy}^2$ \\

4 & 4 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.15cm] (1) at (0,0.866) {1};
    \node[draw,circle,minimum size=0.15cm] (2) at (-0.5,0) {2};
    \node[draw,circle,minimum size=0.15cm] (3) at (0.5,0) {3};
    \draw[-] (1) to [out=-150,in=90] (2);
    \draw[-] (1) to [out=-90,in=30] (2);
    \draw[-] (1) to (2);
    \draw[-] (1) to (3);
\end{tikzpicture} 
& $ f_{x} f_{xxx} f_{yyyy} - f_x f_{yyy} f_{xxxy} + 3 f_{x} f_{xyy} f_{xxyy} - 3 f_x f_{xyyy} f_{xxy} - f_{xxx} f_y f_{xyyy} + f_{xxxx} f_y f_{yyy} - 3 f_y f_{xyy} f_{xxxy} + 3 f_y f_{xxy} f_{xxyy}$ \\

4 & 4 &
\begin{tikzpicture}[baseline=0]
    \node[draw,circle,minimum size=0.15cm] (1) at (0,1) {1};
    \node[draw,circle,minimum size=0.15cm] (2) at (-0.9511,0.3090) {2};
    \node[draw,circle,minimum size=0.15cm] (3) at (-0.5878,-.8090) {3};
    \node[draw,circle,minimum size=0.15cm] (4) at (0.5878,-.8090) {4};
    \node[draw,circle,minimum size=0.15cm] (5) at (0.9511,0.3090) {5};
    \draw[-] (1) to (2);
    \draw[-] (1) to (3);
    \draw[-] (1) to (4);
    \draw[-] (1) to (5);
\end{tikzpicture}
& $\left(f_{x}\right)^{4} f_{yyyy} - 4 \left(f_{x}\right)^{3} f_{y} f_{xyyy} + 6 \left(f_{x}\right)^{2} \left(f_{y}\right)^{2} f_{xxyy} - 4 f_{x} \left(f_{y}\right)^{3} f_{xxxy} + f_{xxxx} \left(f_{y}\right)^{4}$ \\
\end{tabular}
\caption{Details and graphical representations of all partial transvectants of weight $4$ of copies of a function $f$ under a transformation in $A(2)$.}
\label{tab:transvectants2}
\end{table}

Owing to the group-subgroup structure some of these partial transvectants appear in signatures
from other transformation groups, so for future reference, we define:
\begin{equation}
    \label{eq:cde}
    \begin{aligned}
        C &= f_{xx}f_{yy}-f_{xy}^2, \\
        D &= f_x^2f_{yy} - 2f_xf_yf_{xy} + f_{xx}f_y^2, \\
        E &= f_{xxx}f_y^3 - 3f_{xxy}f_xf_y^2 + 3f_{xyy}f_x^2f_y -
        f_{yyy}f_x^3.
    \end{aligned}
\end{equation}

\subsection{Signatures of $SA(2)$ and $A(2)$}

The action of the affine group $A(2)$ arises in images as the motion of
planar objects in images taken from a distant camera. Recall that
because of the subgroup relationship, invariants of $A(2)$ are automatically invariants
of $SA(2)$, but not necessarily vice versa. Any of the partial
transvectants found in Tables~\ref{tab:transvectants1}
and~\ref{tab:transvectants2} are relative invariants of $SA(2)$ (as the determinant
is $1$). Hence, we can form a signature for $SA(2)$ by choosing any combination of
partial transvectants. We choose a signature with three components so that
it can be visualised as a three-dimensional surface; note that this will
not be complete and may well have a very large bad set. For example, we can
choose the function value together with $C$ and $D$ from Eq.~\eqref{eq:cde}
to form the signature:

\begin{equation}
    \label{eq:SA2signature}
    \mathcal{I}_{SA(2)}(f) = \left(f, C, D \right).
\end{equation}

Fig.~\ref{fig:SA2} shows the signature of our test image under an $SA(2)$ transformation.

\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_SA2.png}
\includegraphics[width=.45\textwidth]{Figs/SA2_signature.png}
\caption{A sample $SA(2)$ transformation applied to our test function, and the corresponding signature.}
%\begin{equation*}
%\mathbf{x} \mapsto \frac{1}{1.22}\begin{bmatrix} 0.7 & 1.1 \\ -0.6 & 0.8 \end{bmatrix}\mathbf{x} + \begin{bmatrix} 0.1 \\ -0.2 \end{bmatrix}
%\end{equation*}
\label{fig:SA2}
\end{figure}

For $A(2)$ the weight of the partial transvectants prevents them from being
invariants. There are two options to create a signature. One is to select
ratios of partial transvectants, as discussed in
Section~\ref{sec:Sim2}. It is then beneficial to choose the
numerator and denominator so that they have different critical points,
which can be achieved by adding together partial transvectants of the same
weight as appropriate. Instead, we employ the same normalisation strategy
as previously, projecting the weight four relative invariants $(C^2, D^2,
E)$ onto the unit hypersphere and multiplying by $f$. In this way, our
signature is:
\begin{equation}
    \label{eq:A2signature}
    \mathcal{I}_{A(2)} = \frac{f}{\sqrt{C^4 + D^4 + E^2}}(C^2, D^2, E)
\end{equation}

We will subject our test image to the special affine transformation:
Fig.~\ref{fig:A2} shows the signature of our test image under this transformation.
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_A2.png}
\includegraphics[width=.45\textwidth]{Figs/A2_signature.png}
\caption{A sample $A(2)$ transformation applied to our test function, and the corresponding signature.}
%\begin{equation*}
%\mathbf{x} \mapsto \begin{bmatrix} 0.7 & 1.1 \\ -0.6 & 0.8
%\end{bmatrix}\mathbf{x} + \begin{bmatrix} 0.1 \\ -0.2 \end{bmatrix}
%\end{equation*}
\label{fig:A2}
\end{figure}

\section{Differential invariants through moving frames}\label{sec:movingframes}

So far, the methods that we have seen to construct differential invariants for particular groups are not directly generalisable. It is natural to ask if there is an algorithmic approach that will yield invariants more directly. The answer is a qualified yes. There is a general method, based on Cartan's method of moving frames, that is generally applicable, but finding good solutions with it still requires some ingenuity in general. Olver and co-authors have popularised the approach and made it applicable to object recognition based on outline curves, see~\citet{Olver2005} for an overview, and note particularly the original application paper~\citep{Calabi1998}.  We introduce the method here and then demonstrate its use to find invariants to further groups, but for further technical details see~\citet{OlverCIT}.

A moving frame is a mapping into a group $\mathcal{G}$ that provides a frame of reference along a manifold.  It was extended by~\citet{Cartan35} to the `method of moving frames'. If the group $\mathcal{G}$ is a Lie group then it has a homogeneous space and a frame of reference can be applied on the manifold structure of the group. Considering the actions of such a transformation group enables the identification of invariants to that group since it defines a $\mathcal{G}$-equivariant mapping. A (right-) moving frame is a smooth $\mathcal{G}$-equivariant map $\rho : \mathcal{M} \to \mathcal{G}$ such that $\rho (g \cdot z) = \rho(z) \cdot g^{-1}$, where $z \in J^d$. It is necessary~\citep{Olver2013} for the group act freely. This means that the group action needs to be extended to the entire jet space of derivatives up to order $d$ by implicit differentiation of the group action, a process known as prolongation.

The procedure to construct differential invariants to the action of Lie group $\mathcal{G}$ acting on manifold $\mathcal{M}$ using the method of moving frames is as follows:

\begin{enumerate}
\item Prolong the group action to the jet space of $d$-th order derivatives: $F: J^d \to \mathbb{R}$. 
\item Apply Cartan normalisation:
    \begin{enumerate}
    \item Choose a local cross-section to the group orbits, i.e., a ($d - \dim \mathcal{G}$)-dimensional submanifold $\mathcal{K}$ that intersects transversally at most once with each orbit
    \item $\mathcal{K}$ is specified by $\dim \mathcal{G}$ independent equations $Z_i (z) = c_i$, for $z \in J^d$,  and $Z_i$ scalar-valued functions and $c_i$ constants.
    \item The right-moving frame $g = \rho(z)$ is found by solving the normalisation questions $Z_i (g \cdot z) = c_i$ for the group parameters $g$ (in terms of $z$). In other words, the moving frame is the transformation back to the cross-section.
    \end{enumerate}
\item The differential invariants are not affected by the action of the group: $I(g \cdot z) = I(z) \: \forall z \in \mathrm{dom} I$, or equivalently, if it is constant on the orbits. 
\item Hence $F$ is made invariant by transforming it with a group action: $I(z) = F(g \cdot z) |_{g=\rho(z)}$.  
\end{enumerate}

Cartan normalisation can be considered as identifying a canonical subspace (with some coordinates fixed) in the jet space to which any point in the jet space can be mapped by a group action.

\citet{OlverCIT} uses the group $SE(2)$ on curves as an example, recovering the Euclidean curvature as a second-order invariant. We continue this example to compute invariants to $SE(2)$ for images, before moving on to other planar Lie group examples: $E(2)$, the M\"obius group ($PSL(2,\mathbb{C})$) and the projective group ($PSL(3,\mathbb{R})$). Finally, we apply the method to an infinite-dimensional pseudo-Lie group, the conformal diffeomorphisms.

\subsection{The Special Euclidean Group SE(2)}
\label{sec:SE2}

An element of the special Euclidean group $SE(2)$ acts on an element $(x, y)$ of $\mathbb{R}^2$ to produce $(\bar{x}, \bar{y})$ as:
\begin{equation*}
  \begin{aligned}
    \bar{x} &= x\cos\theta  - y\sin\theta + t_x \\
    \bar{y} &= x\sin\theta  + y\cos\theta + t_y.
  \end{aligned}
\end{equation*}
We prolong the group action to form the jet space $J^2$ with coordinates $(x,
y, f, f_x, f_y, f_{xx}, f_{xy}, f_{yy})$ via the chain rule. The
transformation rule for the derivatives is given by:
\begin{equation}
  \begin{bmatrix}
  \bar f_{\bar x} \\ \bar f_{\bar y} \\ \bar f_{\bar{x}\bar{x}} \\ \bar f_{\bar{x}\bar{y}} \\ \bar f_{\bar{y}\bar{y}}
  \end{bmatrix} = 
  \begin{bmatrix}
 c & -s & 0 & 0 & 0 \\
 s & c & 0 & 0 & 0 \\
0 & 0 & c^2 & -2cs & s^2 \\
0 & 0 & cs & c^2 - s^2 & -cs \\
0 & 0 & s^2 & 2cs & c^2 \\
  \end{bmatrix}
  \begin{bmatrix}
f_x \\ f_y \\ f_{xx} \\ f_{xy} \\ f_{yy} 
  \end{bmatrix},
\label{eqn:SE2}
 \end{equation}
where $c = \cos \theta$ and $s = \sin \theta$. Because the group has three
parameters, we choose the cross-section of the group orbits to be $\bar{x}
= 0, \bar{y} = 0, \bar{f}_{\bar{y}} = 0$. We also require that $\bar{f}_{\bar{x}}
> 0$ to uniquely define the moving frame. Note that the group action is
locally free (away from critical points) when prolonged to $J^1$, however
we prolong to $J^2$ in order to compute invariants. The moving
frame is then given by the prolonged Euclidean transformation that maps an
element of  $J^2$ to this cross-section. It can be thought of as rotating
the image about that point so that its gradient is pointing in the positive
$x$ direction, and then translating the image so the point is mapped to the
origin.

This choice of moving frame gives $\cos\theta = f_x/\lVert\nabla f\rVert$
and $\sin\theta = -f_y/\lVert \nabla f \rVert$. The parameters $t_x, t_y$
are then $t_x = -x\cos\theta + y\sin\theta$ and $t_y = -x\sin\theta -
y\cos\theta$, however these have no effect on the derivatives, and will
henceforth be ignored.

The remaining elements not fixed by the cross-section, $\bar{f}$,
$\bar{f}_{\bar{x}}$, $\bar{f}_{\bar{x}\bar{x}}$,
$\bar{f}_{\bar{x}\bar{y}}$, and $\bar{f}_{\bar{y}\bar{y}}$, are therefore
all invariant. This gives the following set of invariants:
\begin{equation}\label{eq:se2invariants}
\begin{aligned}
  K_0 &= f \\
  K_1 &= f_x^2 + f_y^2 & &(= \bar{f}_{\bar{x}}^2)\\
  K_2 &= f_x^2 f_{xx} + 2 f_x f_y f_{xy} + f_y^2 f_{yy} &&(= K_1\bar{f}_{\bar{x}\bar{x}}) \\
  K_3 &= f_x f_y (f_{yy} - f_{xx}) + f_{xy} (f_x^2 - f_y^2) &&(= K_1\bar{f}_{\bar{x}\bar{y}})\\
  K_4 &= f_y^2 f_{xx} - 2 f_x f_y f_{xy} + f_x^2 f_{yy} &&(= K_1\bar{f}_{\bar{y}\bar{y}}).
\end{aligned}
\end{equation}
Note that we have multiplied each of the second derivative invariants by the
invariant $K_1$ so that they decay smoothly to zero at critical
points.  Interestingly, $K_4$ is the affine invariant $D$ from
Eq.~\eqref{eq:cde} and $K_1$ and $K_2$ are also Euclidean invariants, so we need to
include $K_3$ in our signature: 
\begin{equation}
    \label{eq:SE2signature}
    I_{SE(2)}(f) = (f, K_1, K_3).
\end{equation}
Fig.~\ref{fig:SE2} shows the signature of our test image under a sample
transformation of $SE(2)$.
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_SE2.png}
\includegraphics[width=.45\textwidth]{Figs/SE2_signature.png}
\caption{A sample $SE(2)$ transform and 3D signature.}
\label{fig:SE2}
\end{figure}

\subsection{The Euclidean Group E(2)}
An element of the Euclidean group $E(2)$ maps an element $(x, y)$ of
$\mathbb{R}^2$ to $(\bar{x}, \bar{y})$ as:
\begin{equation*}
  \begin{aligned}
    \bar{x} &=  x\varepsilon\cos\theta  - y\varepsilon\sin\theta + t_x \\
    \bar{y} &= x\sin\theta  + y\cos\theta + t_y.
  \end{aligned}
\end{equation*}
where $\varepsilon \in \{-1, 1\}$ and the remaining parameters are the same
as for $SE(2)$. If $\varepsilon = 1$ it is a rigid transformation, and if
$\varepsilon = -1$ it contains a reflection.

This time, the group action is not free on $J^1$ (with elements of the form
$(x, y, f, f_x, f_y)$) away from critical points because there are two
possible transformations that map to the cross-section $(0, 0, \bar{f},
\bar{f}_{\bar{x}}, 0)$ where $\bar{f}_{\bar{x}} > 0$, one that reflects the
gradient into the positive $x$ direction, and one that rotates.

We again prolong the group action to $J^2$, but this time we expect to need
to use a second derivative term to determine $\varepsilon$ for the moving
frame. The derivative transformations are given by:
\begin{equation}
  \begin{bmatrix}
  \bar f_{\bar x} \\ \bar f_{\bar y} \\ \bar f_{\bar{x}\bar{x}} \\ \bar f_{\bar{x}\bar{y}} \\ \bar f_{\bar{y}\bar{y}}
  \end{bmatrix} = 
  \begin{bmatrix}
 \varepsilon c & -\varepsilon s & 0 & 0 & 0 \\
 s & c & 0 & 0 & 0 \\
0 & 0 & c^2 & -2cs & s^2 \\
0 & 0 & \varepsilon cs & \varepsilon(c^2 - s^2) & -\varepsilon cs \\
0 & 0 & s^2 & 2cs & c^2 \\
  \end{bmatrix}
  \begin{bmatrix}
f_x \\ f_y \\ f_{xx} \\ f_{xy} \\ f_{yy} 
  \end{bmatrix},
\label{eqn:E2prolongation}
\end{equation}
where again $c = \cos\theta$ and $s = \sin\theta$.

We begin by choosing the same cross-section as for $SE(2)$, namely $\bar{x} = 0,
\bar{y} = 0, \bar{f}_{\bar{y}} = 0, \bar{f}_{\bar{x}} > 0$. This gives
$\cos\theta = \varepsilon f_x / \lVert \nabla f \rVert$ and $\sin\theta =
-\varepsilon f_y / \lVert \nabla f \rVert$. Under this transformation,
regardless of whether $\varepsilon = 1$ or $\varepsilon = -1$, the
derivative $\bar{f}_{\bar{x}} = \lVert \nabla f \rVert$. 
The only second derivative term in which $\varepsilon$ appears is
the $\bar{f}_{\bar{x}\bar{y}}$ term, so we use this to resolve the sign. Substituting in for $\cos\theta,
\sin\theta$, we see that:
\begin{equation*}
    \bar{f}_{\bar{x}\bar{y}} = \frac{\varepsilon(f_xf_y(f_{yy} - f_{xx}) + (f_x^2 -
    f_y^2)f_{xy})}{f_x^2 + f_y^2}.
\end{equation*}
This suggests choosing $\varepsilon$ to make $\bar{f}_{\bar{x}\bar{y}} \ge
0$. The moving frame is then fully specified, and the remaining derivatives
give the same invariants $K_0, K_1, K_2, K_4$ as for $SE(2)$,
\eqref{eq:se2invariants}, the one difference being:
\begin{equation}\label{eq:e2invariants}
    \hat{K}_3 = \lvert f_x f_y (f_{yy} - f_{xx}) + f_{xy} (f_x^2 - f_y^2)\rvert
\end{equation}

These invariants are a functionally independent, complete, set of
polynomial differential invariants up to second derivatives. Note that the
invariants found previously by tensor contraction Eq. (\ref{eq:e2tensor}) in
Section~\ref{sec:contraction} can be expressed in terms of them:
\begin{align*}
  f = f &= K_0 \\
  f_if_i = f_x^2 + f_y^2 &= K_1 \\
  f_{ii} = f_{xx} + f_{yy} &= \frac{K_2 + K_4}{K_1} \\
  f_{ij} f_{ij} = f_{xx}^2 + 2f_{xy}^2 + f_{yy}^2 &= \frac{K_2^2 + K_4^2 + 2K_3^2}{K_1^2} \\
  f_i f_j f_{ij} = f_x^2 f_{xx} + 2f_x f_y f_{xy} + f_y^2 f_{yy} &= K_2.
\end{align*}

\subsection{The M\"{o}bius Group}

The M\"obius group $PSL(2,\mathbb{C})$ acts on the Riemann sphere $\overline{\mathbb{C}} = \mathbb{C}\cup\infty$ by:
\begin{equation} 
\phi\colon \overline{\mathbb{C}} \to \overline{\mathbb{C}},\quad \phi(z) = \frac{\alpha z + \beta}{\gamma z + \delta},\quad \alpha,\beta,\gamma,\delta\in\mathbb{C}, \quad
\alpha\delta-\beta\gamma\ne 0.
\end{equation}

Identifying $\R^2$ with $\mathbb{C}$, the M\"obius group is a real 6-dimensional local Lie group acting on $\mathbb{R}^2$. 
It is the smallest nonlinear planar group that contains $SE(2)$, and it also has direct applications in image processing since it arises in the {\em conformal camera} model of vision, in which scenes are projected radially onto a sphere \citep{Lenz1990,Turski2004}; it is also the set of biholomorphic maps of the Riemann sphere. 
%Different shapes may be related by M\"obius transformations, as is explored by Petukhov \cite{petukhov1989non} for biological objects: in one example, the human skull grows approximately by M\"obius transformations.  Note also that Thompson \cite{thompson1942growth} famously deformed images of one species to match those of another, many of his examples using conformal mappings, and thus may be approximated (or even determined by) M\"obius transformations. Discussing this work, Milnor \cite{milnor2010growth}  argued that ``The geometrically simplest way to change the relative size of different body parts would be by a conformal transformation. It seems plausible that this simplest solution will often be the most efficient, so that natural selection tend to choose it.'' (Milnor was thinking of 3D conformal transformations, whose restriction to 2D is the M\"obius transformations.)

We are free to take $\delta=1$. Then $\bar f = f$ means that $z$ is translated to $\beta$; we can therefore take $\beta=0$. This leaves 4 group parameters $a,b,c,d$, where $\alpha = a + i b$, $\gamma = c + i d$, to be determined as follows. We prolong the action to second derivatives of $f$, which is sufficient to allow the group to act freely. There are 5 derivatives of $f$ up to order 2, so there will be at least 1 invariant. In order to find further invariants the prolongation to 3rd derivatives is required, which will result in at least another four invariants. This simple counting argument works because the group acts freely on the prolongation. %There are 9 derivatives of $f$ of order 1, 2, and 3, so there must be at least 5 invariants.

The prolonged action is:
\begin{equation}
\arraycolsep=1.4pt
\left[
\begin{array}{c}
  \bar f_{\bar x} \\ \bar f_{\bar y} \\ \bar f_{\bar x \bar x} \\ \bar f_{\bar x \bar y} \\ \bar f_{\bar y \bar y} 
 \end{array}
 \right]
 = 
\left[
\begin{array}{ccccc}
 a & b & 0 & 0 & 0 \\
 -b & a & 0 & 0 & 0 \\
 2\mu & -2 \rho & a^2 & 2 a b & b^2 \\
 2 \rho & 2\mu & -a b & a^2\!-b^2 & a b \\
 -2\mu & 2 \rho & b^2 & -2 a b & a^2 \\
\end{array}
\right]\left[
\begin{array}{c}
f_x \\ f_y \\ f_{xx} \\ f_{xy} \\ f_{yy} \\
 \end{array}
 \right].
\label{eqn:mobius}
 \end{equation}

\noindent where $\mu = b d - a c$ and $\rho = b c + a d$.

%\begin{equation}
%\arraycolsep=1.4pt
%\left(
%\begin{array}{c}
  %\bar f_x \\ \bar f_y \\ \bar f_{xx} \\ \bar f_{xy} \\ \bar f_{yy} \\ \bar f_{xxx} \\ \bar f_{xxy} \\ \bar f_{xyy} \\ \bar f_{yyy} 
 %\end{array}
 %\right)
 %= 
%\left(
%\begin{array}{ccccccccc}
 %a & b & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 %-b & a & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 %2\mu & -2 \rho & a^2 & 2 a b & b^2 & 0 & 0 & 0 & 0 \\
 %2 \rho & 2\mu & -a b & a^2\!-b^2 & a b & 0 & 0 & 0 & 0 \\
 %-2\mu & 2 \rho & b^2 & -2 a b & a^2 & 0 & 0 & 0 & 0 \\
 %-6 (2 b c d+a \lambda) & 6 (2 a c d-b\lambda) & 6 a \mu &
  %M_{64}  & -6 b \rho & a^3 & 3 a^2 b & 3 a b^2 & b^3 \\
 %-6 (2 a c d-b\lambda) & -6 (2 b c d+a \lambda) & M_{73}& M_{74} & 
 %M_{75} & -a^2 b & a^3\!-2 a b^2 & 2
   %a^2 b\!-b^3 & a b^2 \\
 %6 (2 b c d+a \lambda) & -6 (2 a c d-b\lambda) & M_{83} & M_{84} & M_{85} & a b^2 & b^3\!-2
   %a^2 b & a^3\!-2 a b^2 & a^2 b \\
 %6 (2 a c d-b\lambda) & 6 (2 b c d+a \lambda) & 6 b \mu & 
  %M_{94} & 6 a \rho & -b^3 & 3 a b^2 & -3 a^2 b & a^3 \\
%\end{array}
%\right)\left(
%\begin{array}{c}
%f_x \\ f_y \\ f_{xx} \\ f_{xy} \\ f_{yy} \\ f_{xxx} \\ f_{xxy} \\ f_{xyy} \\ f_{yyy} 
 %\end{array}
 %\right).
%\label{eqn:mobius}
 %\end{equation}
 %where
 %\begin{equation*}
 %\begin{array}{lll}
%M_{64}=-12abc+6d(a^2-b^2), & 
%M_{73}=4 d a^2+6 b c a-2 b^2 d, & 
%M_{74}=-6 c a^2+12 b d a+6 b^2 c, \\
 %M_{75}=-2 d a^2-6 b c a+4 b^2 d, &
 %M_{83}=2 c a^2-6 b d a-4 b^2 c, &
 %M_{84}=6 (d a^2+2 b c a-b^2 d),\\
 %M_{85}=-4 c a^2+6 b d a+2 b^2 c,& 
 %M_{94}= -12a b d + 6c(a^2-b^2),\\
%\lambda=d^2-c^2, & 
%\mu = b d - a c, & 
%\rho = b c + a d.
%\end{array}
%\end{equation*}

The moving frame calculation is then:
\begin{enumerate}
\item The frame $\bar f_{\bar x}=1$, $\bar f_{\bar y}=0$ determines two group parameters:
$$ a = f_x/(f_x^2 + f_y^2),\quad b = f_y/(f_x^2 f_y^2).$$

\item The frame $\bar f_{\bar x \bar y}=\bar f_{\bar y \bar y}=0$ determines the next group parameters:
$$ c = (2 f_xf_yf_{xyy} - f_{xx}f_y^2 - f_{yy}f_x^2)/(2 f_x^2 + f_y^2)^2$$
$$ d = (f_{xy}f_x^2 - f_x f_y f_{xx} - f_{xy}f_y^2 - f_{yy} f_x f_y)/(2 f_x^2 + f_y^2)^2.$$

\item The second-order invariant is then given by $\bar f_{\bar x \bar x}$ using Eq.~(\ref{eqn:mobius}):
$$\bar f_{\bar x \bar x} = \frac{f_{xx}+f_{yy}}{f_x^2 + f_y^2}.$$
\end{enumerate}

%It is clear that there must be 2 syzygies between the relative invariants $J_1,\dots,J_4,K_1,\dots,K_4$.
%They are easily found algebraically to be
%$$K_2 = -K_4 + J_1 J_4$$
%and
%$$K_3=-K_1 +J_1 J_3 + 2 J_1^2 J_2^2.$$

The four third-order invariants can be deduced from the moving frame. They
are rational functions with numerators of degree 6 and denominators $(f_x^2
+ f_y^2)^4$, and are relative invariants of weight 8.  Clearing
denominators so as to work with polynomials, we choose $M_1 = (f_{xx} +
f_{yy})^4$ and two of the four third-order invariants $M_2$ and $M_3$:
\begin{align*} M_2&=
f_y ^5 f_{yyy} + \frac{9}{2} f_y ^2 f_{yy}^2 f_x ^2 + 
 f_y ^3 f_{yyy} f_x ^2 + \frac{3}{2} f_{yy}^2 f_x ^4 - 
 9 f_y ^3 f_{yy} f_x  f_{xy} + 
 3 f_y  f_{yy} f_x ^3 f_{xy} + \frac{3}{2} f_y ^4 f_{xy}^2 \\ \nonumber
 &- 9 f_y ^2 f_x ^2 f_{xy}^2 
+ \frac{3}{2} f_z ^4 f_{xy}^2 + 
 3 f_y ^4 f_x  f_{xyy} + 3 f_y ^2 f_x ^3 f_{xyy} + 
 3 f_y ^4 f_{yy} f_{xx} + 3 f_{yy} f_x ^4 f_{xx} + 
 3 f_y ^3 f_x  f_{xy} f_{xx} \\ \nonumber
&- 9 f_y  f_x ^3 f_{xy} f_{xx} 
+ \frac{3}{2} f_y ^4 f_{xx}^2 + 
 \frac{9}{2} f_y ^2 f_x ^2 f_{xx}^2 + 3 f_y ^3 f_x ^2 f_{xxy} + 
 3 f_y  f_x ^4 f_{xxy} + f_y ^2 f_x ^3 f_{xxx} + 
 f_x ^5 f_{xxx}, 
\end{align*}

\begin{align*}
M_3&=-3 f_y  f_{yy}^2 f_x ^3 + 
 f_y ^2 f_{yyy} f_x ^3 + f_{yyy} f_x ^5 + 
 9 f_y ^2 f_{yy} f_x ^2 f_{xy} - 
 3 f_{yy} f_x ^4 f_{xy} - 6 f_y ^3 f_x  f_{xy}^2 + 
 6 f_y  f_x ^3 f_{xy}^2 \\ \nonumber
&- 3 f_y ^3 f_x ^2 f_{xyy} - 
 3 f_y  f_x ^4 f_{xyy} - 3 f_y ^3 f_{yy} f_x  f_{xx} + 
 3 f_y  f_{yy} f_x ^3 f_{xx} + 3 f_y ^4 f_{xy} f_{xx} - 
 9 f_y ^2 f_x ^2 f_{xy} f_{xx} \\ \nonumber
 &+ 3 f_y ^3 f_x  f_{xx}^2 + 3 f_y ^4 f_x  f_{xxy} + 
 3 f_y ^2 f_x ^3 f_{xxy} - f_y ^5 f_{xxx} - 
 f_y ^3 f_x ^2 f_{xxx}.
\end{align*}
We then form a signature by our usual projection method from the image $f$
and these three weight $8$ relative invariants:
\begin{equation}
    \label{eq:mobiussignature}
    \mathcal{I}_{\text{M\"obius}}(f) = \frac{f}{\sqrt{M_1^2 + M_2^2 +
    M_3^2}} (M_1, M_2, M_3)
\end{equation}
An example is shown in Fig.~\ref{fig:mobius}, although note that the
signature is singular when $f_x=f_y=f_{xx}+f_{yy}=0$, a codimension 1
phenomenon for images. 
\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_Mobius.png}
\includegraphics[width=.45\textwidth]{Figs/Mobius_signature.png}
\caption{A sample M\"obius transform and 3D signature.}
\label{fig:mobius}
\end{figure}

%In previous work \cite{mobius}, two independent 3rd order M\"obius invariants were calculated
%from the classical M\"obius geometry of curves. Namely, the M\"obius arclength 
%of a curve is $\sqrt{|\kappa_s|}ds$, were $ds$ is Euclidean arclength. Under any conformal map $\phi$,
%both $ds$ and $1/\|\nabla f\|$ scale by $|\phi'|$. The sign of $\kappa_s$ is also  invariant,
%hence $\kappa_s/\|\nabla f\|^2$ is  invariant for any invariant curve. 
%Calculating this invariant
%for the level sets of $f$ and for their orthogonal curves yields two invariants of $f$ that
%we call $\lambda_t$ and $\lambda_n$, respectively. Writing 
%them out in full and comparing to the invariants founds above shows that they are related very simply, by

%$$ \frac{K_2}{\|\nabla f\|^8} = \lambda_t$$
%$$\frac{K_4}{\|\nabla f\|^8} =\lambda_n$$

%We do not know of any equivalent characterization of the invariants $K_1$, $K_3$ in
%terms of classical M\"obius geometry.

%We now have that
%$$ (J_1^4, J_2^4, J_1 J_3, J_1 J_4, K_1, K_4)$$
%are 6 relative invariants of weight 8. Projecting any subset to a sphere provides a M\"obius signature.
%For example, the image $f$ together with the projection of $(J_1^4, J_2^4, K_1)$ provides a 3-dimensional signature.
%Any such signature that includes $J_1$ and $J_2$ is

\subsection{The Projective Group}

The action of the projective group arises naturally through the movements of the camera viewing planar objects.
We consider the eight-dimensional group $PSL(3,\R)$ acting on $\R^2$ by projective transformations, i.e., 
\begin{equation}
(x,y) \mapsto \left(\frac{a x + b y + j}{l + g x + h y}, \frac{c x + d y + k}{l+ g x + h y}\right).
\end{equation}

Requiring that $\bar f = f$ enables us to clear the translation coefficients, so $j=k=0$; for convenience we will also take $l=1$. This leaves 6 group parameters $(a,b,c,d,g,h)$ to be determined. For free action it is necessary to prolong the group action to 3rd derivatives of $f$. Since there are 9 derivatives of $f$ of order 1, 2, and 3, there will be at least 3 invariants.

The prolonged action is:
\begin{equation}
\arraycolsep=1.4pt
\label{eq:prolong}
%\rotatebox{90}{$
\left[
\begin{array}{c}
  \bar f_{\bar x} \\ \bar f_{\bar y} \\ \bar f_{\bar x \bar x} \\ \bar f_{\bar x \bar y} \\ \bar f_{\bar y \bar y} \\ \bar f_{\bar x \bar x \bar x} \\ \bar f_{\bar x \bar x \bar y} \\ \bar f_{\bar x \bar y \bar y} \\ \bar f_{ \bar y \bar y \bar y} 
 \end{array}
 \right]
 = 
\left[
\begin{array}{ccccccccc}
 a & c & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 b & d & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 -2 a g & -2 c g & a^2 & 2 a c & c^2 & 0 & 0 & 0 & 0 \\
 -b g-a h & -d g-c h & a b & b c+a d & c d & 0 & 0 & 0 & 0 \\
 -2 b h & -2 d h & b^2 & 2 b d & d^2 & 0 & 0 & 0 & 0 \\
 6 a g^2 & 6 c g^2 & -6 a^2 g & -12 a c g & -6 c^2 g & a^3 & 3 a^2 c & 3 a c^2 & c^3 \\
 2 g \alpha_1 & 2 g \beta_1 & -2 a \alpha_2 & -4 \delta_1 & -2 c \beta_2
   & a^2 b & a \gamma_1 & c \gamma_1 & c^2 d \\
 2 h \alpha_2 & 2 h \beta_2 & -2 b \alpha_1 & -4 \delta_2 & -2 d \beta_1
   & a b^2 & b \gamma_1 & d \gamma_2 & c d^2 \\
 6 b h^2 & 6 d h^2 & -6 b^2 h & -12 b d h & -6 d^2 h & b^3 & 3 b^2 d & 3 b d^2 & d^3 \\
\end{array}
\right]\left[
\begin{array}{c}
f_x \\ f_y \\ f_{xx} \\ f_{xy} \\ f_{yy} \\ f_{xxx} \\ f_{xxy} \\ f_{xyy} \\ f_{yyy} 
 \end{array}
 \right].%$}
 \end{equation}

\noindent where $\alpha_1 = b g + 2 a h, \alpha_2 = 2 b g + a h$, $\beta_1 = d g + 2 c h, \beta_2 = 2 d g + c h$, $\gamma_1 = b c + 2 a d, \gamma_2= 2 b c + a d$, and $\delta_1 = (b c g+a d g+a c h), \delta_2 = (b d g+b c h+a d h)$.

We begin the construction of the moving frame by choosing $\bar f_{\bar x}
= 1$ and $\bar f_{\bar y}=  0$, which determines the group parameters $a =
(1-c f_y)/f_x.$ and $b = -d f_y/f_x$. Direction substition then shows that
$\bar f_{\bar y \bar y} = D d^2 / f_x^2$, where $D = f_{x}^2 f_{yy} + f_y^2
f_{xx} - 2 f_x f_y f_{xyy}$ was listed in Eq.~\eqref{eq:cde} as an invariant of
$A(2)$, which is a subgroup of $PSL(3,\R)$. Note that $D$ and
$\bar f_{\bar y \bar y}$ have the same sign. If $D < 0$, we choose the
frame $\bar{f}_{\bar y \bar y} = -1$ and if $D \ge 0$ we choose $\bar
f_{\bar y \bar y} = 1$. Then $d = f_x/|D|^{1/2}.$

The frame $\bar f_{ \bar y \bar y \bar y} = 0$ determines the group
parameter $ h = \frac{f_{yyy}f_x^3 - 3 f_x^2 f_y f_{xyy} + 3 f_x f_y^2
f_{xxy} - f_y^3 f_{xxx}}{6|D|^{3/2}},$ while $\bar f_{ \bar x \bar y 
y}=0$ provides $ c = \frac{h f_x^2}{d D} - \frac{f_x f_{xy} - f_y
f_{xx}}{D}$, and $\bar f_{ \bar x \bar x}=0$ gives $ g = \frac{1}{2
f_x^2}(f_{xx} + 2 c ( f_x f_{xy} - f_y f_{xx} ) + c^2 D).$ In fact, the
numerator of the expression for $h$ was another invariant of $A(2)$ labelled as
$E$ in Eq.~\eqref{eq:cde}.

%\begin{equation}
%E := f_{yyy}f_x^3 - 3 f_{xyy}f_x f_y^2 + 3 f_{xxy}f_x^2 f_y - f_{xxx}f_y^3.
%\end{equation}

Having determined the frame, any function of the remaining derivatives $\bar f_{ \bar x \bar x \bar x}$, $\bar f_{ \bar x \bar x \bar y}$, and $\bar f_{ \bar y \bar y \bar y}$, as given in Eq.~(\ref{eq:prolong}), provides invariants. We choose: 
\begin{equation*}
 (\bar f_{\bar x\bar x\bar x}, \bar f_{\bar x\bar x\bar y}^2, \bar f_{\bar
 x\bar y\bar
 y})=\left(\frac{J_1}{D^6},\frac{J_2}{D^9},\frac{J_3}{D^3}\right).
\end{equation*} 
The terms $J_1, J_2, J_3$ can be written in terms of $D$ and $E$, together
with a polynomial of degree $n$ denoted by $P_n$. 
\begin{equation*}
    J_1 = E^4 + D P_{13},\quad J_2 = (E^3 + D P_9)^2,\quad J_3=E^2 -12 D P_5.
\end{equation*}

The $J_i$ are polynomials of degree 16, 24, and 8, respectively, and are extremely complicated when written out explicitly. They do have one benefit, though, which is that they  all have denominators given by integer powers of $D$, resolving the ambiguity caused by the sign of $D$. Therefore, $ (D^{18}, J_1^3, J_2^2, J_3^6)$ are all relative invariants of weight 36, and so projecting any subset of them of them to a sphere yields a third-order invariant signature.

$J$ vanishes on generic images: $J=0$ when $D=E=0$, which is a codimension 0 phenomenon for images. In particular, $D=E=0$ at critical points. However, these signatures can be simplified and made more robust. Their structure suggests considering the combination $J_1 - J_3^2$, which obeys:
$$J_1 - J_3^2 = 144 D^4 P_4.$$

As $J_1/D^6$ and $J_3/D^3$ are invariant, so is $P_4/D^2$. 
Moreover, the numerator of this new invariant does not vanish at critical points.  It has the structure:
$$P_4 = -4(\det f_{ij})^2 + Q_4,$$
where $Q_4$ is a polynomial of degree 4 that vanishes when $f_x=f_y=0$. Hence, the relative invariant signature of weight 12 $ (D^6, P_4^3, (E^2-12D P_5)^2)$ can be projected to $S^2$, yielding a projective invariant of images that is singular only when $D=E=P_4=0$, which is a codimension 1 phenomenon, since  $D=E=0\Rightarrow P_4=0$. In particular, at critical points it tends to $(0, (2\det f_{ij})^6, 0),$ and hence critical points with $\det f_{ij}\ne 0$ have signature value $(0,1,0)$.
Including $J_2$ provides a 4-dimensional signature set, but does not change the codimension of the set of bad images.

%For completeness we write out the invariant signature we use explicitly:

%$Q_4=4 f_x  f_{xy} ^2 f_{xyy}  - 
%f_x ^2 f_{xyy} ^2 - 
 %2 f_{yyy}  f_x  f_{xy}  f_{xx}  + 
 %2 f_{yy}  f_x  f_{xyy}  f_{xx}  - 
 %6 f_{y}  f_{xy}  f_{xyy}  f_{xx}  + 
 %2 f_{y}  f_{yyy}  f_{xx} ^2 + 
 %f_{yyy}  f_x ^2 f_{xxy}  - 
 %6 f_{yy}  f_x  f_{xy}  f_{xxy}  + 
 %4 f_{y}  f_{xy} ^2 f_{xxy}  + 
 %f_{y}  f_x  f_{xyy}  f_{xxy}  + 
 %2 f_{y}  f_{yy}  f_{xx}  f_{xxy}  - 
 %f_{y} ^2 f_{xxy} ^2 + 
 %2 f_{yy} ^2 f_x  f_{xxx}  - 
 %f_{y}  f_{yyy}  f_x  f_{xxx}  - 
 %2 f_{y}  f_{yy}  f_{xy}  f_{xxx}  + f_{y} ^2 f_{xyy}  f_{xxx}, $

%$P_5=f_{yyy}  f_{x} ^3 f_{xy}  - 
%f_{yy}  f_{x} ^2 f_{xy} ^2 + 
 %2 f_{y}  f_{x}  f_{xy} ^3 - 
 %f_{yy}  f_{x} ^3 f_{xyy}  - 
 %f_{y}  f_{x} ^2 f_{xy}  f_{xyy}  + 
 %f_{yy} ^2 f_{x} ^2 f_{xx}  - 
 %f_{y}  f_{yyy}  f_{x} ^2 f_{xx}  - 
 %2 f_{y}  f_{yy}  f_{x}  f_{xy}  f_{xx}  - 
 %f_{y} ^2 f_{xy} ^2 f_{xx}  + 
 %2 f_{y} ^2 f_{x}  f_{xyy}  f_{xx}  + 
 %f_{y} ^2 f_{yy}  f_{xx} ^2 + 
 %2 f_{y}  f_{yy}  f_{x} ^2 f_{xxy}  - 
 %f_{y} ^2 f_{x}  f_{xy}  f_{xxy}  - 
 %f_{y} ^3 f_{xx}  f_{xxy}  - 
 %f_{y} ^2 f_{yy}  f_{x}  f_{xxx}  + 
 %f_{y} ^3 f_{xy}  f_{xxx}. $

%Further invariants can be found by noting that, for any planar transformation group and any two invariants
%$I^1$ and $I^2$, $\nabla I^1 \times \nabla I^2$ is a relative invariant of weight 1; that is:
%$$\nabla I^1 \times \nabla I^2\mapsto (\det \phi')(\nabla I^1 \times \nabla I^2).$$
%Applying this operation to $f$ and the four third-order invariants yields $5\cdot4/2=10$ fourth-order invariants.

\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{Figs/f_transformed_PSL3R.png}
\includegraphics[width=.45\textwidth]{Figs/PSL3R_signature.png}
\caption{A sample $PSL(3,\mathbb{R})$ transform and 3D signature.}
\label{fig:PSL3R}
\end{figure}

\subsection{Conformal Diffeomorphisms $\mathbf{Diff}_{\mathbf{con}}$}

The conformal camera model is well known. The group of conformal diffeomorphisms is the subgroup of diffeomorphisms (i.e., smooth functions with smooth inverses) that are angle-preserving:

\begin{equation}
\angle (u,v) = \angle (D \phi(x) . u, D \phi(x).v),
\end{equation}

\noindent where $\angle(u, v)$ denotes the angle between tangent vectors
$u$ and $v$ based at $x \in \mathbb{R}^2$, $\phi \in \textrm{Diff}$ and $D$
represents the Jacobian. The group is infinite-dimensional, but still small
enough that images have local differential invariants.

The conformal diffeomorphisms provides a particularly nice example of the
moving frame method. Because the group is infinite dimensional, we need to
prolong the group action to the infinite jet space $J^{\infty}$ with
elements of the form $(x, y, f, f_x, f_y, f_{xx}, f_{xy}, f_{yy}, \ldots)$.
%Recall the moving frame produces a conformal map $\psi$ for each element of
%$J^\infty$ that maps it to a suitably chosen cross-section.

It is convenient to work in the complex variable $z = x + iy$. Without loss of
generality, we can assume $x = y = 0$. We begin the construction of our
cross-section by choosing $\bar{x} = \bar{y} = 0$. Then $\psi^{-1}$ 
can be represented locally by a Taylor series of the form:
\begin{equation*}
    z = \psi^{-1}(\bar{z}) = c_1 \bar{z} + c_2 \bar{z}^2 + c_3\bar{z}^3 + \cdots,
\end{equation*}
where $c_k = a_k + ib_k, k = 1, 2, \ldots$. We solve for the parameters
$a_k, b_k$ in stages.  To find $c_1 = a_1 + i b_1$ we need two constraints.
Differentiating $f(x, y) = \bar{f}(\bar{x}, \bar{y})$ and substituting in
our existing cross-section constraints $\bar{x} = \bar{y} = 0$ gives:
\begin{align*}
    \bar{f}_{\bar{x}} = f_x x_{\bar{x}} + f_y y_{\bar{x}} &= a_1 f_x + b_1
    f_y, \\
    \bar{f}_{\bar{y}} = f_x x_{\bar{y}} + f_y y_{\bar{y}} &= -b_1 f_x + a_1
    f_y. 
\end{align*}
Note that $\bar{x} = \bar{y} = 0$ removes all higher order derivatives from
the cross-section equation. Adding two extra constraints to the cross
section, $\bar{f}_{\bar{x}} = 1, \bar{f}_{\bar{y}} = 0$, this system is readily
solved to give:
\begin{equation*}
    \begin{aligned}
        a_1 &= f_x / (f_x^2 + f_y^2), &  b_1 &= f_y / (f_x^2 + f_y^2).
    \end{aligned}
\end{equation*}
We then repeat the process, this time using the second derivative
computations. There are two new parameters, $a_2, b_2$, to solve
for, but there are three derivatives. We use the cross-section equations
$\bar{f}_{\bar{x}\bar{x}} = 0, \bar{f}_{\bar{x}\bar{y}} = 0$ to solve for
$a_2$ and $b_2$, the remaining derivative $\bar{f}_{\bar{y}\bar{y}}$ is then the
sole second-order invariant, which we also saw for the M\"obius group: 
\begin{equation}
    \bar{f}_{\bar{y}\bar{y}} = \frac{f_{xx} + f_{yy}}{f_x^2 + f_y^2} =
    \frac{C_1}{f_x^2 + f_y^2}.
\label{eqn:C1}
\end{equation}
Continuing in this way, at derivative order $n$ two real group parameters
enter: $a_n$ and $b_n$. There are $n+1$ independent derivatives of $f$ of
order $n$, so there must be $n-1$ new independent invariants at each order.
Hence at third order there are two more invariants, using the constraints
$\bar{f}_{\bar{x}\bar{x}\bar{x}} = 0, \bar{f}_{\bar{x}\bar{x}\bar{y}} = 0$,
the remaining derivatives are invariant:
\begin{align}
    \bar{f}_{\bar{x}\bar{y}\bar{y}} &= 
    \begin{multlined}[t]
        \frac{1}{(f_x^2 + f_y^2)^3} \left[
f_y^3f_{yyy} + f_x^2f_yf_{yyy}
            -2f_y^2f_{yy}^2 - 2f_{xx}f_y^2f_{yy} -
            4f_xf_{xy}f_yf_{yy} - 2f_x^2 f_{xx}f_{yy} \right. \\ 
+ \left. f_{xxy}f_y^3 +
            f_x^2 f_{xxy}f_y - 4 f_x f_{xy} f_{xx} f_y + f_x^3 f_{xxx} -
    2f_x^2 f_{xx}^2 + f_x^3 f_{xyy} \right]
    = \frac{C_2}{(f_x^2 +
f_y^2)^3}\end{multlined} \\
    \bar{f}_{\bar{y}\bar{y}\bar{y}} &= 
    \begin{multlined}[t]
        \frac{1}{(f_x^2 + f_y^2)^3} \left[
f_xf_y^2f_{yyy} +
        f_x^3f_{yyy} - 2f_x f_y f_{yy}^2 + 2f_{xy}f_y^2f_{yy} -
    2f_x^2f_{xy}f_{yy} - f_{xxx}f_y^3 \right. \\ 
-  f_{xyy}f_y^3 + f_xf_{xxy}f_y^2 +
2f_{xy}f_{xx}f_y^2 - f_x^2 f_{xxx} f_y + 2f_xf_{xx}^2 f_y - f_x^2 f_{xyy}
f_y \\\left.+ f_x^3 f_{xxy} - 2f_x^2 f_{xy} f_{xx}\right]
= \frac{C_3}{(f_x^2 + f_y^2)^3}
\end{multlined}\label{eqn:C3}
\end{align}
Note that the denominators of all three of these equations are the same up
to index. We therefore clear denominators as usual and form the
signature using our usual projection technique:
\begin{equation}
    \label{eq:conformalsignature}
\mathcal{I}_{\text{Diff}_\text{con}}(f) = \frac{f}{\sqrt{C_1^6 + C_2^2 +
C_3^2}}(C_1^3, C_2, C_3)
\end{equation}
This signature is continuous at all $(x,y)$ such that the three quantities
are not simultaneously equal to zero, which occurs when $f_x = f_y = f_{xx}
= f_{yy} = 0$, which is a codimension 1 phenomenon for images.

An example for the conformal map $\varphi$ defined through $\varphi^{-1}$ by
$\varphi^{-1}(z) = \frac{3}{8}(z - 2)^2 - \frac{3}{2}$ is shown in
Fig.~\ref{fig:Diffcon}.
\begin{figure}
\centering 
\includegraphics[width=.45\textwidth]{Figs/f_transformed_Diffcon.png}
\includegraphics[width=.45\textwidth]{Figs/Diffcon_signature.png}
\caption{A sample $\text{Diff}_{con}$ Transform and 3D signature.}
\label{fig:Diffcon}
\end{figure}

%However, we first note that the M\"obius action is locally identical to the full conformal group up to and including second order. Therefore, there are no new conformal invariants, apart from the M\"obius ones, up to second order. It is possible to identify some invariants via direct calculation using the defining geometric properties of the conformal group.

%At second order, we first consider the action of conformal maps on the Euclidean invariants $f_i f_i$ and $f_{ii}$.

%Let $\psi := \varphi^{-1}$. For any diffeomorphism, the derivatives of $f$ transform as
%$$f_i \mapsto f_j \psi^j_i$$
%and
%$$f_{ij} \mapsto f_{kl} \psi^k_i\psi^l_j + f_{k}\psi^k_{ij}.$$
%Therefore the Laplacian of $f$ transforms as
%$$ f_{ii} \mapsto f_{kl} \psi^k_i \psi^l_i + f_k \psi^k_{ii}.$$
%When $\psi=u + i v$ is conformal, $\psi_{k,ii}=0$ because the components of conformal maps are harmonic functions,
%and the Cauchy-Riemann equations give $\psi_{k,i}\psi_{l,i} = (u_x^2 + u_y^2)\delta_{kl}$. Therefore,
%$$ f_{ii} \mapsto f_{ii}(u_x^2+u_y^2).$$
%Here $u_x^2+u_y^2=\det\psi^k_i$ is the local scaling factor of the transformation.
%Note also that
%$$f_i f_i \mapsto f_k f_l \psi_{k,i} \psi_{l,i} = f_k f_k (u_x^2 + u_y^2).$$
%That is, $f_{ii}$ and $f_i f_i$ are both relative invariants of weight 2, while $f_{ii}/(f_i f_i)$ is a conformal invariant, as we saw for the M\"obius group. By itself, it is not useful for images as it is singular at critical points of $f$.

%The transformation of 1st derivatives is more clearly written in complex variables as $$ f_x + i f_y \mapsto a_1 (f_x + i f_y).$$ Given any two real invariants $g,h$, the real and imaginary parts of $(g_x+i g_y)(h_x-i h_y)$ are relative invariants of weight 2. This generates the required two invariants of order 3. (We have repeated and confirmed this calculation using the method of moving frames.)

%Specifically, let $J_1 = f_i f_i$ and  $J_2 = f_{ii}$, so that  $g = J_2/J_1$ is invariant.  We want to work with polynomials, so we clear denominators and let $J_3+iJ_4 = J_1^2 (g_x + i g_y)  (f_x - i f_y)$. $J_3$ and $J_4$ are a differential polynomials of $f$ of degree 4 and order 3, and are relative invariants of weight 6. They can be written explicitly as functions of the basic invariants of $SE(2)$ as
%$$J_3 = f_k f_k f_{iij}f_j - 2 f_{ii} f_{jk} f_j f_k$$
%and (writing $a_i\times b_i$ for $a_1 b_2 - b_2 a_1$)
%$$ J_4 = f_k f_k f_{iij}\times f_j - 2 f_{ii} f_{jk} f_k\times f_j.$$
%Under conformal maps we have
%$$ (J_1,J_2,J_3,J_4) \mapsto (|a_1|^2 J_1, |a_1|^2 J_2, |a_1|^6 J_3,|a_1|^6 J_4).$$

%Therefore, projecting $(J_1^3,J_2^3,J_3)$ to the unit sphere, together with $f$ itself, provides a conformal signature set in $\R\times S^2$.  It is continuous at all $(x,y)$ such that $J_1$, $J_2$, and $J_3$ are not simultaneously equal to zero.  Now $J_1=0$ implies $f_1=f_2=0$, and $J_2=0$ implies $f_{ii}=0$, but these together imply $J_3=0$. Thus the signature set is discontinuous when $f_1=f_2=f_{ii}=0$, a codimension-1 phenomenon for images.

%Similarly, projecting $(J_1^3,J_2^3,J_3,J_4)$ to the unit sphere, together with $f$ itself, provides a conformal signature set in $\R\times S^3$. This is also singular when $f_1=f_2=f_{ii}=0$. These singularities are not removable. A typical image near a `bad' point takes the form $f = a(x^2-y^2) + b x y + c x^3 + d x^2 y + e x y^2 + g y^3$.  For such an $f$, $(J_1^3,J_2^3,J_3,J_4)=\mathcal{O}(x^6,x^3,x^3,x^3)$.  The cubic forms always change sign and do not generically lead to a well-defined direction as $(x,y)\to(0,0)$.

\subsection{$\mathrm{Diff}_{\mathrm{vol}}$ and $\mathrm{Diff}$}

Two other infinite-dimensional groups are of interest: the volume-preserving diffeomorphisms, which act as the space of transformations of an incompressible fluid, and the full group of planar diffeomorphisms, which are widely-used in image registration~\citep{YounesBook}. However, in these cases there are no differential invariant signatures of greyscale images.

The volume-preserving diffeomorphisms $\mathrm{Diff}_{\mathrm{vol}}$ are those diffeomorphisms that preserve the volume form (area form in 2D). They can be defined by one function of two variables, the generating function $S$; one possibility is $S(x',y')$ where $(x,y)\mapsto(x',y')$ according to:
$$ x'=\frac{\partial S}{\partial y'},\quad y = \frac{\partial S}{\partial x}.$$

Consider expanding $S$ in Taylor series. The linear terms in the map are determined from the quadratic terms in $S$, which contain 3 group parameters, and act on 2 coordinates, $f_x$ and $f_y$.  The quadratic terms in the map are determined from the cubic terms in $S$, which contain 4 group parameters, and act on 3 coordinates, $f_{xx}$, $f_{xy}$, and $f_{yy}$. At order $n$ we have $n+2$ new group parameters acting on $n+1$ coordinates. Thus, barring some exceptional behaviour in the action, we do not expect any invariants at generic points. This can be confirmed by considering a small image patch with $f(x,y)=x.$ Let $v=\psi_y\partial/\partial_x - \psi_x\partial/\partial_y$ be the generator of an area-preserving map. The image $f$ transforms to $x + \varepsilon \psi_y(x,y)$ which (by choosing $\psi$ appropriately) can be any function close to $x$.
At nongeneric points, there can be invariants. For example, the area enclosed in any level set of $f$ is invariant. At minima or maxima of $f$, this will lead to invariants expressed in terms of the derivatives of $f$.

In the full diffeomorphism group $\mathrm{Diff}$ there are no differential invariants using 1 or even 2 colours (colour will be discussed in the Section~\ref{sec:colour}). For example, with $k=2$, for generic $f^1$, $f^2$ we can choose a diffeomorphism (i.e., choose local coordinates) so that $f^1=x$ and $f^2=y$.

\section{Discussion}\label{sec:colour}

%\todo[inline]{Number of derivatives}
%\todo[inline]{Noise; integral invariants, Fourier}
%\todo[inline]{Is there a general theory - syzygy??}
%\todo[inline]{Difficulties of a practical implementation}

The recognition of objects within images as their appearance changes based on camera motion remains an unsolved problem. In this paper we have used three different methods of computing differential invariant signatures to identify appropriate signatures for images transformed under the planar Lie groups that are of interest for this question. The resulting three-dimensional signatures are not always complete, but they are relatively fast to compute. However, there are some issues that arise with turning this into a practical method.

The first is a standard issue for images: producing a computationally robust method of numerically approximating derivatives and hence invariants is not trivial, particularly for the higher-orders, where noise and rounding errors can dominate. There are three primary considerations: noise and texture; flat areas, where the derivatives are all identically zero; and the fact that conventional numerical differentiation is not well-posed~\citep{Florack1993}. Some considerations of these points and some potential approaches to solving them for practical implementations are considered by~\citet{Calabi1998} and~\citet{Florack1993}, but the question of how to create a full and effective method of computing numerical derivatives for images remains open. Practical approaches to using the signatures identified in this paper for real-world images will be considered in future work.

The second problem is that there are hidden symmetries within the lattice of groups we consider. For example, the relation $SA(2)\subset \mathrm{Diff}_{\mathrm{vol}}$ acts an obstruction to $SA(2)$, since many of the invariants of $SA(2)$ are also invariant under the larger group. While there will be more invariants for the smaller group, care needs to be taken to check that the set chosen are unique to that group. The same issue arises with $A(2)$, even though $A(2)$ is not even a subgroup of $\mathrm{Diff}_{\mathrm{vol}}$.

It might be hoped that colour information can help. The formulation we provided in the introduction considered $k$ channels of information, but so far we have considered only greyscale images ($k=1$). Most real images are three-channel colour (RBG), and multi-spectral images can have hundreds of channels of measurement. It is therefore natural to consider whether this information is useful for the computation of invariants, in particular by allowing lower-order derivatives to form a suitable signatures. However, in practice the situation is more complicated that it first appears because the information in the channels is not independent: colour is built by the intensities in the three channels, and so there is correlation between them. %We briefly sketch some possibilities for some of the groups that we have considered earlier, but note that the correlation means that the invariant signatures identified next are na\"ive, and should be viewed as best-case scenarios. 
%for $k$ colours, signature sets of dimension up to $k + k(k+1)/2$ can be chosen from the invariants $f^i$ and $f^i_m f^j_m$ for $i,j=1,\dots, k$.

For $E(2)$, for $k\ge 2$, first derivatives suffice, but for both $SA(2)$ and $A(2)$ the hidden symmetry noted above mean that there are no signatures with only first derivatives.

The picture is more positive for the infinite-dimensional groups.  For $\mathrm{Diff}_{\mathrm{con}}$, first-order signature sets are possible for $k \ge 2$ colours. They are formed from the relative invariants $(f^j_x + i f^j_y)(f^m_x - i f^m_y)$ of weight 2 (where $f^j$ is the $j$the colour channel). We illustrate this using the Hopf fibration for $k=2$. The weight 2 relative invariant $$J = (2(f^1_x + i f^1_y)(f^2_x - i f^2_y), ||\nabla f^1\|^2 - \|\nabla f^2\|^2)$$ obeys $\|J\|^2 = \|\nabla f^1\|^2 + \|\nabla f^2\|^2$, hence projecting to the unit sphere shows that $(f^1,f^2,J/\|J\|)$ provides a 4-dimensional signature set that is singular only when $\nabla f^1 = \nabla f^2 = 0$, a codimension 2 phenomenon for images. For $\mathrm{Diff}_{\mathrm{vol}}$ extra channels mean that different invariants now exist: there are $k(k-1)/2$ first-order invariants $ \nabla f^i \times \nabla f^j$. Then the three-dimensional signature $ (f^1,f^2, \nabla f^1 \times \nabla f^2)$ determines (generically and locally) the image up to an area-preserving map. We can choose the area-preserving map so that $f^1=x$. The remaining freedom is of the area-preserving maps that preserve $x$: these are the shears $y\mapsto y + g(x)$. The signature for each fixed $x$ now gives $(f^2,f^2_y)$. This is the standard signature curve for the group of translations in $y$. Thus, $f^2$ is determined by the signature up to a translation in $y$, i.e., an area-preserving map.

In the full diffeomorphism group $\mathrm{Diff}$ there are no differential invariants using 1 or 2 colours. For example, with $k=2$, for generic $f^1$, $f^2$ we can choose a diffeomorphism (i.e., choose local coordinates) so that $f^1=x$ and $f^2=y$. We therefore take $k\ge 3$. This provides the first-order relative invariants of weight 1 $\nabla f^i\times \nabla f^j$, and hence a first-order signature. However, a zeroth-order signature set is possible: $(f^1,\dots,f^k)$, the image of the image. It is (locally and generically) complete. We can choose coordinates so that $f^1=x$ and $f^2=y$; then the signature locally determines $f^3,\dots,f^k$ as functions of $f^1$ and $f^2$, i.e., the image.

%\begin{description}
%\item[$E(2)$]
%Considering $E(2)$, for $k\ge 2$, first derivatives suffice. However, note that for $k\ge 3$ the three-dimensional signature $(f^1,f^2,f^3)$ is invariant under all local planar diffeomorphisms and so a higher-dimensional signature, involving at least first derivatives, is required.
%\item[$SA(2)$]
%For $SA(2)$, one can take transvectants of any components of $f$ to get invariants. For $k=2$ colours, there is only one  invariant formed from first derivatives, namely $((f^1)',(f^2)')_x = f^1_x f^2_y - f^1_y f^2_x$ (the Poisson bracket $\{f^1,f^2\}$). So again two derivatives are necessary, even for a 3-dimensional signature.

%For $k\ge 3$ colours there are plenty of invariants formed from first derivatives, namely $\nabla f^i \times \nabla f^j$ for $i<j$. However, these are all invariant under the bigger group $\mathrm{Diff}_{\mathrm{vol}}$ of local volume-preserving diffeomorphisms, and so the Poisson brackets cannot form a complete invariant for $SA(2)$: two derivatives are always needed no matter how many colours we have. 
%The relation $SA(2)\subset \mathrm{Diff}_{\mathrm{vol}}$ acts an obstruction to $SA(2)$, an instance of the phenomenon of hidden symmetry.
%\item[$A(2)$]
%The same problem occurs with $A(2)$. There are plenty of invariants formed from two derivatives: already 9 when $k=2$. However, as with $SA(2)$, we can never get away with one derivative because all the first-derivative invariants, such as $\{f^1,f^2\}/\{f^1,f^3\}$, are also invariant under $\mathrm{Diff}_{\mathrm{vol}}$ (even though $A(2)$ is not even a subgroup of $\mathrm{Diff}_{\mathrm{vo}l}$).
%\item[$PSL(2,\mathbb{C})$]
%Recall that M\"obius and conformal transformations cannot be distinguished by their action on second derivatives. Therefore at least three derivatives are needed, no matter how many colours we have. We can use, for example, the invariants arising from the curvature of the level sets of each colour. Replacing the relative invariant $\|\nabla f\|^2$ by $\sum_{j=1}^k \|\nabla f^j\|^2$, and $f_{ii}$ by $\sum_{j=1}^k (f^j_{ii})^2$, increases the codimension of the singular images from 1 to $3k-2$.
%\item[$PSL(3,\mathbb{R})$]
%In $PSL(3, \mathbb{R})$ for $2$-colour images, there are three second-order relative invariants of weight 2, namely $(D_1,D_2,(\nabla f^1\times \nabla f^2)^2)$. Together with $(f^1,f^2)$, these yield a 4-dimensional $\mathbb{R}^2\times S^2$-valued signature using only second derivatives. With $k>2$ colours there are many more second-order invariants; however, as the action of the projective group on first derivatives is indistinguishable from that of the full diffeomorphism group, it is not possible to have a first-order signature that distinguishes projectively-related images from diffeomorphically-related ones.
%\item[$\mathrm{Diff}_{\mathrm{con}}$]
%For the infinite-dimensional group $\mathrm{Diff}_{\mathrm{con}}$, first-order signature sets are possible for $k \ge 2$ colours. They are formed from the relative invariants $(f^j_1 + i f^j_2)(f^m_1 - i f^m_2)$ of weight 2. We illustrate this using the Hopf fibration for $k=2$. The weight 2 relative invariant $$J = (2(f^1_1 + i f^1_2)(f^2_1 - i f^2_2), ||\nabla f^1\|^2 - \|\nabla f^2\|^2)$$ obeys $\|J\|^2 = \|\nabla f^1\|^2 + \|\nabla f^2\|^2$, hence projecting to the unit sphere shows that $(f^1,f^2,J/\|J\|)$ provides a 4-dimensional signature set that is singular only when $\nabla f^1 = \nabla f^2 = 0$, a codimension 2 phenomenon for images.
%\end{description}
%When considering colour images, for $k\ge 2$ in $\mathrm{Diff}_{\mathrm{vol}}$ there are $k(k-1)/2$ first-order invariants $ \nabla f^i \times \nabla f^j$.  We consider the case $k=2$ and the 3-dimensional signature $$ (f^1,f^2, \nabla f^1 \times \nabla f^2).$$ This signature determines (generically and locally) the image up to an area-preserving map.  We have seen that we can generically choose the area-preserving map so that $f^1=x$, and we assume that this has been done. The remaining freedom is that of the area-preserving maps that preserve $x$: these are the shears $y\mapsto y + g(x)$. The signature for each fixed $x$ now gives $(f^2,f^2_y)$. This is the standard signature curve for the group of translations in $y$. Thus, $f^2$ is determined by the signature up to a translation in $y$, i.e., an area-preserving map.
%In the full diffeomorphism group $\mathrm{Diff}$ there are no differential invariants using 1 or 2 colours. For example, with $k=2$, for generic $f^1$, $f^2$ we can choose a diffeomorphism (i.e., choose local coordinates) so that $f^1=x$ and $f^2=y$. We therefore take $k\ge 3$. This provides the first-order relative invariants of weight 1 $\nabla f^i\times \nabla f^j$, and hence a first-order signature. However, a zeroth-order signature set is possible: $(f^1,\dots,f^k)$. This is the image of the image. It is (locally and generically) complete. We can choose coordinates so that $f^1=x$ and $f^2=y$; then the signature locally determines $f^3,\dots,f^k$ as functions of $f^1$ and $f^2$, i.e., the image.

Putting all of this information together, Table~\ref{tab:derivs} summarises one of the most important parameters of differential invariant signatures for the transformation groups that we have considered, namely the highest order of derivative of the image that is needed. None of the groups need more than third-order differentiation, and usually less, especially for colour images.

%even this can be problematic for real images. %Our aim in this paper has been to explore the mathematical theory required to compute the invariants.

\begin{table}[h]
\begin{center}
\begin{tabular}{| l | c c c |}
\hline
 & \multicolumn{3}{|c|}{Degree} \\
$\mathcal{G}$ & $k=1$ & $k=2$ & $k=3$ \\
\hline
$SE(2)$ & 2 &1 & 1 \\
$E(2)$ & 2 & 1 & 1 \\
$Sim(2)$ & 2 & 1 & 1 \\
$SA(2)$ & 2 & 2 & 2 \\
$A(2)$ & 3 & 2 & 2 \\
$PSL(2,\mathbb{C})$ & 3 & 3 & 3 \\
$PSL(3,\R)$ & 3 & 2 & 2 \\
$\mathrm{Diff}_{\mathrm{vol}}$ & -- & 1 & 1 \\
$\mathrm{Diff}_{\mathrm{con}}$ & 3 & 1 & 1 \\
$\mathrm{Diff}$ & -- & -- & 0 \\
\hline
\end{tabular}
\caption{Degree (order of highest derivative of $f$) needed to construct a differential invariant signature of $f$ as a function of the group $\mathcal{G}$ and the number of colours $k$ of $f$. The entry `--' indicates that there is no differential invariant signature in that case.}
\label{tab:derivs}
\end{center}
\end{table}

In future work we will seek the underlying mathematical structure that we have not fully exposed. For example, all of the second-order differential invariants of $E(2)$ play known roles in geometric analysis:

\begin{itemize}
\item $f_x^2 + f_y^2=\|\nabla f\|^2$ is the Lagrangian density for the Laplacian;
\item $f_{xx} + f_{yy}$ is the Laplacian, which plays a key role in Euclidean and conformal geometry;
\item $f_{xx}f_x^2 + 2 f_{xy} f_x f_y + f_{yy} f_y^2 $ is the `infinity-Laplacian'; 
\item $f_{xx}f_x^2 + 2 f_{xy} f_x f_y + f_{yy} f_y^2 $ is also a Lagrangian density for $-2(f_{xx} f_{yy} - f_{xy}^2)$, which arises in the Monge--Amp\`ere equations and can be written in terms of the final second-order invariant as $-2 (f_{xx} f_{yy} - f_{xy}^2) = 2 f_{xy}^2 - f_{xx}^2 - f_{yy}^2$
\end{itemize}

Functions of these invariants are invariants or relative invariants for many groups that contain $E(2)$, such as those that we have considered in this paper.

Function $D$ from Eq.~\eqref{eq:cde}, which appeared in $SE(2)$, $SA(2)$ and $PSL(3,\mathbb{R})$, is also known as the Bateman equation~\citep{Andriopoulos}, and is linked to the Hodograph transformation~\citep{Rosenhaus}.

\bibliographystyle{plainnat}
\bibliography{invariants}

\end{document}