### 2021-03-02 paper/letter.tex
As submitted

\documentclass[12pt,a4paper]{letter}
\usepackage[utf8]{inputenc}
%\usepackage[a4paper,left=2.5cm, right=2.5cm, top=2cm, bottom=2.5cm]{geometry}
\usepackage[a4paper,left=2cm, right=2cm,height=26cm]{geometry}
\usepackage{graphicx}
\usepackage{url}

\begin{document}
\thispagestyle{empty}

\textbf{To}
\\
The Editor
\\
SIAM Journal on Imaging Sciences
\\

{\bf Subject}:  Submission of Manuscript

Manuscript title: Differential invariant signatures for planar Lie group transformations of images \\
Authors: Richard Brown, Stephen Marsland, and Robert McLachlan

\vspace{2em}

Dear Editor,

we write to submit the above research paper for consideration by your journal. The paper considers the practicality of differential invariant signatures for image recognition under different transformations, some of which relate to camera models. It extends the mathematical research on invariant signatures for Lie groups, provides a tutorial example of how to use methods such as moving frames to derive invariants, provides examples for a series of important groups, and considers the practical questions that would need to be answered to make a fully useful application. We believe that readers of SIIMS would find it of interest.

Please note that to get the full pdf under 5MB in size we have had to shrink the images. Higher-quality versions of them are available.

Some possible referees that we believe are knowledgeable about this area include:

\begin{itemize}
\item Peter Olver, U Minnesota,  \url{olver@umn.edu}
\item Cheri Shakiban, University of St Thomas, St Paul, \url{cshakiban@stthomas.edu}
\item Alain Trouv\'{e}, ENS-Cachan \url{ trouve@cmla.ens-cachan.fr}
\item Irina Kogan, North Carloin State University, \url{iakogan@math.ncsu.edu}
\item Allen Tannenbaum, Stony Brook University, \url{arobertan@cs.stonybrook.edu}
\end{itemize}

\closing{We look forward to hearing from you,
\\ Professor Stephen Marsland\\Victoria University of Wellington, New Zealand\\ \url{stephen.marsland@vuw.ac.nz}}

\end{document}